### 2024-10-04 study2.qmd
updates

---
title: "Study 2 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
---

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,haven,tibble,ggplot2,gt,ggh4x,lme4)

s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
s2 <- s2 |> mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id)

s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  relocate(refClass,state1,state2, .after = id)

s2_b1 <- s2 |> select(id,state1,refClass,TENPERCENT,ROUNDED, ACFAMILY1:TOTALAPTWO1) |> 
  mutate( 
    block = 1,
    state=state1,
  pct_goal=case_when (TENPERCENT == 1 ~ "10%", TENPERCENT == 2 ~ "15%"), 
  rounded = case_when(ROUNDED == 1 ~ "Rounded", ROUNDED == 2 ~ "Not Rounded")
  ) |> relocate(id,block,refClass,pct_goal,rounded,state)  |> 
  select(-TENPERCENT,-ROUNDED, -state1)

s2_b2 <- s2 |> select(id,state2,refClass,TENPERCENT,ROUNDED, ACFAMILY2:TOTALAPTWO2) |> 
  mutate( 
    block = 2,
    state=state2,
  pct_goal=case_when (TENPERCENT == 1 ~ "15%", TENPERCENT == 2 ~ "10%"), 
  rounded = case_when(ROUNDED == 1 ~ "Not Rounded", ROUNDED == 2 ~ "Rounded")
  ) |> relocate(id,block,refClass,pct_goal,rounded,state) |> 
  select(-TENPERCENT,-ROUNDED, -state2)

# pivot data into longer format, so that each appliance type is on its own row, so for each appliance we have the family value, state, value, plan 1 value, and plan 2 value
# so for the first row for id 113029, we would have ACFAMILY1 ACSTATE1 ACAPONE1 ACAPTWO1, then the next row would be HEATFAMILY1 HEATSTATE1 HEATAPONE1 HEATAPTWO1, etc.
# so each of the bill sources, AC, HEAT, WATER, REFRIG, OTHER, will now be on its own row, with the family, state, value, plan 1 value, and plan 2 value

s2_b1_long <- s2_b1 |>
  pivot_longer(
    cols = c(ACFAMILY1:TOTALAPTWO1),
    names_to = c("appliance", "measure"),
    names_pattern = "(AC|HEAT|WATER|REFRIG|OTHER|TOTAL)(FAMILY1|STATE1|APONE1|APTWO1)",
    values_to = "value"
  ) |>
  pivot_wider(
    names_from = measure,
    values_from = value
  ) |>
  rename(
    family = FAMILY1,
    state_avg = STATE1,
    plan1 = APONE1,
    plan2 = APTWO1
  ) |>
  relocate(appliance, .after = refClass)

s2_b2_long <- s2_b2 |>
  pivot_longer(
    cols = c(ACFAMILY2:TOTALAPTWO2),
    names_to = c("appliance", "measure"),
    names_pattern = "(AC|HEAT|WATER|REFRIG|OTHER|TOTAL)(FAMILY2|STATE2|APONE2|APTWO2)",
    values_to = "value"
  ) |>
  pivot_wider(
    names_from = measure,
    values_from = value
  ) |>
  rename(
    family = FAMILY2,
    state_avg = STATE2,
    plan1 = APONE2,
    plan2 = APTWO2
  ) |>
  relocate(appliance, .after = refClass)

s2_long <- bind_rows(s2_b1_long, s2_b2_long) |> 
  arrange(id, block)

# pivot so that plan1 and plan2 are on their own rows
s2_long <- s2_long |>
  group_by(id,block) |> 
  pivot_longer(
    cols = c(plan1, plan2),
    names_to = "plan",
    values_to = "value"
  ) |> 
  mutate(change = value-family, state_dif = value-state_avg,plan=as_factor(plan))

#write.csv(s2_long,"data/study_2_tg.csv")

# compute total kWh reduction, and total % reduction 
sum1 <- s2_long |> 
  filter(appliance !="TOTAL") |> 
  group_by(id,refClass,state,pct_goal,rounded,block,plan) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif))

```

## Inspect response distribution for outliers

```{r}

outliers <- sum1 |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>2.5)

outlier_id <- outliers |> filter(is_outlier) |> pull(id) |> unique()

outlier_summary <- outliers %>%
  group_by(pct_goal, rounded, refClass,state) %>%
  summarise(
    num_outliers = sum(is_outlier, na.rm = TRUE),
    total = n(),
    outlier_percentage = (num_outliers / total) * 100
  )

outlier_plot <- ggplot(outliers, aes(x = state, y = z_score, color = is_outlier)) +
  geom_jitter(alpha = 0.5) +
  facet_wrap(~ pct_goal) +
  labs(
    title = "Z-Scores of Change in Energy Usage by State - outliers colored",
    x = "State",
    y = "Z-Score",
    color = "Outlier"
  ) +
  theme_minimal()

# Assess skewness and normality of the data
library(e1071)
# Calculate skewness and kurtosis for the `pct_change` variable
skewness_value <- skewness(sum1$pct_change, na.rm = TRUE)
kurtosis_value <- kurtosis(sum1$pct_change, na.rm = TRUE)

cat("Skewness:", skewness_value, "\n")
cat("Kurtosis:", kurtosis_value, "\n")

# Shapiro-Wilk test for normality (sample size limitation)
shapiro_test <- shapiro.test(sum1$pct_change[1:5000])  # Using first 5000 observations if large dataset
print(shapiro_test)
```

::: column-page-right

## Before and after excluding outliers

```{r}
#| layout-ncol: 1
#| fig.width: 10
#| fig.height: 8

#outlier_plot

sum1 |> ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=state), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  theme_minimal() +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

# plot densities in a better way to account for the positive skew

sum1 <- sum1 |> 
  filter(!id %in% outlier_id)

sum1 |> 
  filter(!id %in% outlier_id) |>
  ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=state), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  theme_minimal() +
  labs(title="After outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

```

## Individual Participant responses

- Facets indicate the Reference class (%, kWh, USD) and the goal % reduction (10%, 15%)
- dashed lines also indicate goal percentage - dots on the line indicate participants who met the goal

```{r}
#| fig.width: 11
#| fig.height: 9

  # add geom_vline with x intercept based on pct_goal of the data (which is currently a string of form 10%)
sum1 |> 
  filter(block==1) |>
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=pct_change)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)

```

## Interactive data table

```{r}
# plot distribution of reductions by applicance
# s2_long |> ggplot(aes(x = change)) + geom_histogram(bins = 30) + facet_wrap(~appliance, scales = "free")
sum1 |> mutate(refClass=factor(refClass),pct_goal=factor(pct_goal),rounded=factor(rounded),state=factor(state)) |>
DT::datatable(extensions = 'Buttons',
  options = list(
    dom = 'Blfrtip',
    buttons = list('copy', 'csv', 'excel', 'pdf', 'print'),
    pageLength = 7,
    autoWidth = TRUE
  ),
  filter = 'top') # Add filtering at the top of each column)
#DT::datatable(sum1 |> group_by(id,block) |> summarise(diff(total_kWh),diff(orig_kWh)))
```
:::

## Main Effects and Interactions

**Click on plots to enlarge**

```{r}
#| layout-ncol: 2

sum1 |> ggplot(aes(x = refClass, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  labs(
    title = "Effect of Reference Class",
    x = "Reference Class",
    y = "% Change",
    fill = "Reference Class"
  ) +
  theme_minimal()

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = pct_goal)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  labs(
    title = "Effect of Goal % Reduction",
    x = "Goal % Reduction",
    y = "% Change",
    fill = "Goal % Reduction"
  ) +
  theme_minimal()

sum1 |> ggplot(aes(x = rounded, y = pct_change, fill = rounded)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  labs(
    title = "Effect of Rounding",
    x = "Rounding",
    y = "% Change",
    fill = "Rounding"
  ) +
  theme_minimal()

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  labs(
    title = "Effect of Goal % Reduction and Reference Class",
    x = "Goal % Reduction",
    y = "% Change",
    fill= "Reference Class"
  ) +
  theme_minimal()

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + 
  facet_wrap(~state) +
  labs(
    title = "Effect of Goal % Reduction and Reference Class - Separated by State",
    x = "Goal % Reduction",
    y = "% Change",
    fill= "Reference Class"
  ) +
  theme_minimal()

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = rounded)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + 
  facet_wrap(~rounded) +
  labs(
    title = "Effect of Goal % Reduction and Rounding",
    x = "Goal % Reduction",
    y = "% Change",
    fill= "Rounding"
  ) +
  theme_minimal()

```

```{r}
#| fig.height: 11
#| 
# sum1 |> 
#   ungroup() |> 
#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
#          block=factor(block)) |>
#   group_by(refClass,block,pct_goal) |>
#   summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
#   ggplot(aes(x=mean_pct_change,y=refClass,col=block)) + 
#   geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
#   ggforce::facet_col(~pct_goal)
# 
# 
# sum1 |> 
#   ungroup() |> 
#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
#          block=factor(block)) |>
#   group_by(refClass,,pct_goal,state) |>
#   summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
#   ggplot(aes(x=mean_pct_change,y=refClass,col=pct_goal)) + 
#   geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
#   ggforce::facet_col(~state)
# 
# 
# sum1 |> 
#   ungroup() |> 
#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
#          block=factor(block)) |>
#   group_by(refClass,,pct_goal,state,goal_pct) |>
#   summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
#   ggplot(aes(x=mean_pct_change,y=refClass,col=state)) + 
#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
#   geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
#   ggforce::facet_col(~pct_goal)

  s2_long |> 
    filter(appliance != "TOTAL") |>
  group_by(refClass,appliance,state) |>
  summarise(mean_kWh_change = mean(change),se_pct_change = sd(change)/sqrt(n())) |>
  ggplot(aes(x=mean_kWh_change,y=refClass,col=appliance)) + 
  geom_pointrange(aes(xmin=mean_kWh_change-se_pct_change,xmax=mean_kWh_change+se_pct_change),position = position_dodge(width=0.5)) +
  ggforce::facet_col(~state)

s2_long |> 
  filter(appliance != "TOTAL") |>
  ggplot(aes(x=refClass,y=change,fill=state))+
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  scale_y_reverse() +
  facet_wrap(~appliance)

```

```{r}

# mixed effects model testing effect of refClass, state, and pct_goal on pct_change
# marginal effects plots
library(ggeffects)

m1 <- lmer(pct_change ~ refClass  + (1+plan|id) + (1|state), data = sum1)
summary(m1)
ggpredict(m1, c("refClass")) |> plot()

m1 <- lmer(pct_change ~ pct_goal  + (1|id) + (1|state), data = sum1)
summary(m1)
ggpredict(m1, c("pct_goal")) |> plot()

m1 <- lmer(pct_change ~ pct_goal*rounded  + (1|id) + (1|state), data = sum1)
summary(m1)
ggpredict(m1, c("pct_goal","rounded")) |> plot()

m1 <- lmer(pct_change ~ refClass*pct_goal*rounded  + (1|id) + (1|state), data = sum1)
ggpredict(m1, c("refClass","pct_goal","rounded")) |> plot()

m1 <- lmer(pct_change ~ state*pct_goal*rounded  + (1|id), data = sum1)
summary(m1)
ggpredict(m1, c("state","pct_goal","rounded")) |> plot()

# m1 <- lmer(pct_change ~ refClass*pct_goal*rounded*state  + (1|id), data = sum1)
# ggpredict(m1, c("refClass","pct_goal","rounded","state")) |> plot()

```

```{r}
#| eval: false

s2_long |> 
  filter(appliance != "TOTAL") |>
  ggplot(aes(x=refClass,y=state_dif,fill=state))+
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  scale_y_reverse() +
  facet_wrap(~appliance)

s2_long |> 
  filter(appliance != "TOTAL") |>
  ggplot(aes(x=state_dif,y=value,col=state))+
  geom_point() +
  facet_wrap(~appliance)

sum1 |> 
  ggplot(aes(x=state_dif,y=pct_change,col=state)) +
  geom_point()

# compute total kWh reduction, and total % reduction 
s2_long |> group_by(id,refClass,state,pct_goal,rounded,block,plan) |> 
  filter(appliance !="TOTAL") |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), total_pct = sum(value)/sum(family)) |>
  ggplot(aes(x = pct_goal, y = total_pct, fill = refClass)) + 
  geom_boxplot(position = "dodge") +
  facet_wrap(~state)

s2_long |> group_by(id,refClass,state,pct_goal,rounded) |> 
  filter(appliance !="TOTAL") |> 
  summarise(total_kWh = sum(reduction), total_pct = sum(reduction)/sum(family)) |>
  ggplot(aes(x = pct_goal, y = total_pct, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")

s2_long |> group_by(id,refClass,state,pct_goal,rounded) |> 
  filter(appliance !="TOTAL") |> 
  summarise(total_kWh = sum(reduction), orig_kWh=sum(family), total_pct = round(sum(reduction)/sum(family))) |>
  ggplot(aes(x = state1, y = total_pct, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  facet_wrap(~pct_goal)

s2_long |> group_by(id,refClass,state,pct_goal,rounded,plan) |> 
  filter(appliance !="TOTAL") |> 
  summarise(total_kWh = sum(reduction), total_pct = sum(reduction)/sum(family)) |>
  ggplot(aes(x = refClass, y = total_pct, fill = pct_goal)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  facet_wrap(plan~rounded)

# sum1 |> 
#   ungroup() |> 
#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
#   mutate(id=reorder(id,pct_change)) |> 
#   ggplot(aes(y=id,x=pct_change)) + 
#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
#   ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)

# sum1 |> 
#   ungroup() |> 
#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
#   mutate(id=reorder(id,pct_change)) |> 
#   ggplot(aes(y=id,x=pct_change,col=refClass)) + 
#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
#   ggh4x::facet_nested_wrap(block~pct_goal,axes="all",scales="free",ncol=2) 
# 
# 
# sum1 |> 
#   ungroup() |> 
#     mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
#   ggplot(aes(x=state,y=pct_change,col=state)) + 
#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
#     geom_hline(aes(yintercept=goal_pct),linetype="dashed",alpha=.5) +
#   ggdist::stat_halfeye()+
#   ggh4x::facet_nested_wrap(block~pct_goal,axes="all",ncol=2)

# s2_long <- s2 |>
#   pivot_longer(
#     cols = c(ACFAMILY1:TOTALAPTWO1),
#     names_to = c(".value", "appliance"),
#     names_pattern = "(.*)(FAMILY1|STATE1|APONE1|APTWO1)$",
#     values_to = "value"
#   )

# s2 |> pivot_longer(cols = contains("FAMILY"), names_to = "Family", values_to = "kWh") |> relocate(Family,kWh) |> 
#   pivot_longer(cols = contains("STATE1"), names_to = "State", values_to = "kWh") |> relocate(State)

# 
# s2 |> 
#   pivot_longer(cols=ACFAMILY1:TOTALAPTWO1, names_to = "appliance", values_to = "family") |> relocate(appliance,family)

```