### 2024-10-18 study1.qmd
study 1 analysis scripts

---
title: "Study 1 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
---

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,haven,tibble,ggplot2,gt,ggh4x,lme4,flextable)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

```

```{r}
#| eval: FALSE

# s1 <- haven::read_sav(here::here("data/data_qualtrics/Frequency & Probability - 9-23-18_October 6, 2024_20.39.sav")) # length(unique(s1$ResponseId)) = 588
# qd <- unique(s1$ResponseId)

s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))

#gd <- unique(s1$ResponseId)
# gd %in% qd
# qd %in% gd

s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  rename(id=ResponseId) |>
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

s1 <- s1 |> rename(edu=DEM04) |>
  mutate(edu = factor(case_when(
    edu == 1 ~ "Some schooling, but no diploma or degree",
    edu == 2 ~ "Highschool diploma or GED",
    edu == 3 ~ "Some college",
    edu == 4 ~ "College degree",
    edu == 5 ~ "Some graduate school",
    edu == 6 ~ "Graduate degree",
    edu == 7 ~ "Choose not to answer",
    TRUE ~ NA_character_
  ))) |> relocate(edu, .after = id)

# colnames(s1)
# colnames(s1)[1000:1500]
# colnames(s1)[1500:1843]

# select all columns that start with "A"
# s12 <- s1 |> select(id,state1,state2,refClass,calc,edu,starts_with("A")) |> 
#   select(-starts_with("AIND"))

# 
# t <- s12 |> filter(state1=="Texas") |> 
#   # select only columns that contain "TEX" in the name
#   select(id,state1,state2,refClass,calc,edu,contains("TEX")) 
# 
# t2 <- s12 |> filter(state2=="Texas") |> 
#   select(id,state1,state2,refClass,calc,edu,contains("TEX"))
#

trim_na_and_empty_columns <- function(df) {
  df %>%
    select(where(~!(all(is.na(.)) | (is.character(.) && all(trimws(.) == ""))))) |> 
    select(-starts_with("AIND"))
}

# Create a list of trimmed datasets for each unique ID
dAll <- s1 %>%
  group_by(id) %>%
  group_split() %>%
  map(trim_na_and_empty_columns) %>%
  set_names(unique(s1$id))

# modalities <- c(A = "kWh", D = "USD", P = "Percentage")
# states <- c(CAL = "California", MASS = "Massachusetts", COL = "Colorado", TEX = "Texas")

parse_action_plan <- function(df) {

  action_plan_cols <- df %>% 
  select(-matches("01$|02$|03$|04$|05$|06$|33$|34$|35$|36$|37$")) %>%
    select(
      matches("^[ADP]\\d+(CAL|MASS|COL|TEX)\\d{2,3}$"), #,  # Select action plan columns
      -matches("32$")                                     # Exclude columns ending with '32'
    ) %>%
    names()

  df_long <- df %>%
  select(-matches("01$|02$|03$|04$|05$|06$")) %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value)

  df_long <- df_long |> 
  mutate(
    appliance = case_when(
      item_num %in% c("07","08","09",7:10) ~ "Cooling",
      item_num %in% 11:14 ~ "Heating",
      item_num %in% 15:18 ~ "Water Heating",
      item_num %in% 19:22 ~ "Refrigerator",
      item_num %in% 23:26 ~ "Other Appliance",
      item_num %in% 27:30 ~ "Total kWh",
      item_num == 31 ~ "Preferred Plan",
      item_num %in% 33:37 ~ "Reduction Choices",
      TRUE ~ NA_character_
    )
  ) |> 
  # item_type - family, state, plan1 or plan2
  mutate(item_type=case_when(
    item_num %in% c("07","11","15","19","23","27") ~ "family",
    item_num %in% c("08","12","16","20","24","28") ~ "state_avg",
    item_num %in% c("09","13","17","21","25","29") ~ "plan1",
    item_num %in% c("10","14","18","22","26","30") ~ "plan2",
    TRUE ~ "other"
  ))

# pivot family and state_avg into their own separate columns, plan1 and plan2 into a combined column called plan
df_long <- df_long |> 
  filter(item_type !="other") |> 
  select(-item_num) |> 
  pivot_wider(names_from = c("item_type"), values_from = value) |> 
  pivot_longer(cols = c("plan1","plan2"), names_to = "plan", values_to = "value")

}

test <- dAll |> # map each item in dAll into parse function
  map(parse_action_plan)

s1 <- bind_rows(test)

s1 <- s1 |> group_by(id) |> 
  mutate(trial=row_number(), block=ifelse(trial>=12,2,1), plan=as_factor(plan)) |> 
  select(-trial)

saveRDS(s1, here::here("data/s1_processed.rds"))

```

```{r}
s1 <- readRDS(here::here("data/s1_processed.rds"))

s1 <- s1 |> mutate(change = value-family, state_dif = value-state_avg)

sum1 <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
  mutate(matched_goal = (pct_change == .15))

```

```{r}

sum1 |> ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=state), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  #facet_wrap(~refClass) +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

# descriptive table to do with plot
sum1 |> 
  group_by(refClass,state) |> 
  summarise(mean_pct_change=median(pct_change), sd_pct_change=sd(pct_change), n=n()) |> 
  gt() |> 
  tab_header(title="Descriptive Statistics of Percent Change by State and Reference Class")

```

```{r}
#| fig.width: 10
#| fig.height: 9

  # add geom_vline with x intercept based on pct_goal of the data (which is currently a string of form 10%)
sum1 |> 
  filter(block==1) |>
  ungroup() |> 
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=pct_change)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=.15),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(~refClass,axes="all",scales="free",ncol=2)  + 
  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)))
```

## Looking at % of subjects who hit the target reduction

```{r}

# Step 2: Summarize the percentage of participants meeting the goal for each condition
summary_pct <- sum1 |>
  group_by(refClass) |>
  summarise(
    matched_count = sum(matched_goal),
    total_count = n(),
    pct_matched = matched_count / total_count * 100,
    .groups = "drop"
  )

# Step 3: Create plots for the percentage of subjects in each condition who matched the target pct
ggplot(summary_pct, aes(x = refClass, y = pct_matched, fill = refClass)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Percentage of Participants Matching Target Pct by Condition",
    x = "Reference Class",
    y = "Percentage Matched (%)",
    fill = "Rounding"
  ) +
  theme_minimal()

summary_pct |>
  gt() |>
  tab_header(
    title = md("**Percentage of Participants Matching Target Pct by Condition**"),
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal"
  ) |>
  fmt_number(
    columns = c(pct_matched),
    decimals = 2
  ) |>
  cols_label(
    refClass = "Reference Class",
    rounded = "Rounding",
    pct_goal = "Pct Goal",
    matched_count = "Matched Count",
    total_count = "Total Count",
    pct_matched = "Percentage Matched (%)"
  ) |>
  cols_align(
    align = "center",
    columns = everything()
  ) |>
  tab_spanner(
    label = "Participants",
    columns = c(matched_count, total_count, pct_matched)
  )

```