### 2024-10-24 results.qmd
update with results page

---
title: "Results"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: results_hugo.md
  gfm:
    echo: true
    output-file: results_gfm.md
---

```{r}

library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
options(brms.backend="cmdstanr",mc.cores=4)

walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds"))

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
  mutate(matched_goal = (pct_change == pct_goal))

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

s2_long <- readRDS(here::here("data/s2_processed.rds"))

s2_agg <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
         matched_goal2 = abs(pct_change-pct)<.03)

s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

s2_agg2 <- s2_agg |> group_by(id,refClass,rounded,pct_goal,pct,calc) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

```

## Experiment 1

```{r}
##| label: tbl-s1-bb
##| tbl-cap: "E1. beta-binomial regression results

s1_bb1 <- brm(
  mg | trials(4) ~ refClass,
  family = beta_binomial(),
  cores=4,
  iter=5000,
  data = s1_agg4,
  file=paste0(here::here("data/model_cache",'s1_bb1.rds'))
)

mted1 <- as.data.frame(describe_posterior(s1_bb1, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted1) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")

mted1 |>
  mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
                \(x) plogis(x))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(booktabs = TRUE)

```

```{r}
#| label: fig-s1-mg1
#| fig-cap: Study 1
#| fig-width: 11
#| fig-height: 8

plot(conditional_effects(s1_bb1, effects="refClass"),points=FALSE)

posterior_predict(s1_bb1,ndraws=200) |> array_branch(margin=1) |> 
  map_dfr(
    function(yrep_iter){
      s1_agg4 %>% ungroup() |>  mutate(pred=yrep_iter)
    },
    .id='iter'
  ) |> mutate(iter=as.numeric(iter)) |> 
  ggplot(aes(pred, group=iter)) +
  # Scale density to match the count scale
  geom_line(stat='density', aes(y=..count..), 
            alpha=0.7, color='blue') +
  geom_bar(data=s1_agg4, aes(mg, fill=refClass),
           inherit.aes=FALSE) +
  facet_wrap(~refClass)
```

```{r}
##| label: tbl-s2-bb
##| tbl-cap: "Study 2. beta-binomial regression results
##| 
##|

s2_bb2_r_g <- brm(
  mg | trials(2) ~ refClass*rounded*pct_goal,
  family = beta_binomial(),
  data = s2_agg2,
  cores=4,
  iter=5000,
  file=paste0(here::here("data/model_cache",'s2_bb2_r_g.rds'))
)

mted2 <- as.data.frame(describe_posterior(s2_bb2_r_g, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted2) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")

mted2 |>
  mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
                \(x) plogis(x))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(booktabs = TRUE)
```

```{r}
#| label: fig-s2-mg1
#| fig-cap: Study 2
#| fig-width: 11
#| fig-height: 9

# plot(conditional_effects(s2_bb2_r_g, 
#                          effects = "rounded:pct_goal", 
#                          conditions=make_conditions(s2_bb2_r_g,"refClass" )),
#      points=FALSE,plot=TRUE)

patchwork::wrap_plots(plot(conditional_effects(s2_bb2_r_g),points=FALSE,plot=FALSE))

```

```{r}
#| label: fig-s2-mg2
#| fig-cap: Study 2
#| fig-width: 11
#| fig-height: 10

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = refClass, fill = refClass)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) 
}

pe3td <- s2_agg2 |>  ggplot(aes(x =refClass, y = mg,fill=pct_goal)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
    facet_wrap(~rounded,ncol=1) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation From Target")

pe3ce <- s2_bb2_r_g |> emmeans( ~refClass *rounded*pct_goal, type="response") |>
  gather_emmeans_draws() |>
 condEffects(pct_goal) + labs(y="Absolute Deviation From Band", x="Band Type") + 
 facet_wrap(~refClass,ncol=1)

s2_bb2_r_g |> emmeans(pairwise ~ refClass * rounded, type="response") |> 
  pluck("contrasts") |>
  gather_emmeans_draws() |> 
  group_by(contrast,.draw) |> summarise(value=mean(.value), n=n()) |> 
   ggplot(aes(x=value,y=contrast,fill=contrast)) +
  stat_halfeye() + 
  geom_vline(xintercept=0, alpha=.5) +
  labs(x="Model Error Difference",y="Contrast") + theme(legend.position="none")

```