### 2024-11-04 misc/discrep.qmd
s1 discrepancy checks; index page reset

---
title: "Discrep_Patterns"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: show
code-tools: true
execute:
  warning: false
---

## Study 1

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,tibble, haven,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds"))
s2_long <- readRDS(here::here("data/s2_processed.rds"))

```

```{r}

s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(distinct_state_avg = n_distinct(state_avg)) %>%
  ungroup() %>%
  filter(distinct_state_avg > 1) %>%
  group_by(appliance, state_avg, state, refClass) |> arrange(id)

s1 |> group_by(state,appliance,refClass) |> 
  mutate(state_avg=as.numeric(state_avg)) |> 
  mutate(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  filter(n_distinct_family>1 | n_distinct_state_avg>1)

s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(
    distinct_families = n_distinct(family),
    distinct_state_avgs = n_distinct(state_avg),
  ) %>%
  ungroup() %>%
  filter(distinct_families > 1 | distinct_state_avgs>1) %>%
  group_by(appliance, family, state, refClass) %>%
  summarise(n = n(), .groups = "drop") |> gt()

```

```{r}

state_code_map <- c(
  "CAL" = "California",
  "TEX" = "Texas",
  "COL" = "Colorado",
  "MASS" = "Massachusetts"
)

state_to_code <- state_code_map
code_to_state <- setNames(names(state_code_map), state_code_map)
lookup_code <- function(state) {
  if (state %in% names(state_to_code)) {
    return(state_to_code[[state]])
  } else if (state %in% names(code_to_state)) {
    return(code_to_state[[state]])
  } else {
    return(NA)  # Return NA if not found in either mapping
  }
}

correct_values <- tribble(
  ~state, ~appliance, ~family, ~state_avg,
  # Texas values
  "Texas", "Cooling", 6573, 4249,
  "Texas", "Heating", 6118, 5099,
  "Texas", "Water Heating", 5257, 4396,
  "Texas", "Refrigerator", 2639, 1318,
  "Texas", "Other Appliance", 8459, 7883,
  "Texas", "Total kWh", 29046, 22945,

  # California values
  "California", "Cooling", 2581, 1289,
  "California", "Heating", 6157, 5597,
  "California", "Water Heating", 5061, 4601,
  "California", "Refrigerator", 1266, 1055,
  "California", "Other Appliance", 7608, 6916,
  "California", "Total kWh", 22673, 19458,

  # Colorado values
  "Colorado", "Cooling", 697, 498,
  "Colorado", "Heating", 18052, 16411,
  "Colorado", "Water Heating", 11667, 5832,
  "Colorado", "Refrigerator", 1370, 1142,
  "Colorado", "Other Appliance", 7982, 6652,
  "Colorado", "Total kWh", 39768, 30535,

  # Massachusetts values
  "Massachusetts", "Cooling", 419, 322,
  "Massachusetts", "Heating", 26751, 19108,
  "Massachusetts", "Water Heating", 10543, 5070,
  "Massachusetts", "Refrigerator", 1230, 1025,
  "Massachusetts", "Other Appliance", 7350, 6682,
  "Massachusetts", "Total kWh", 46293, 32207
) |> 
  rowwise() |>
  mutate(state_abrv=lookup_code(state))

 s1 |> 
       filter(#appliance != "Total kWh", 
       plan == "plan1") %>%
        group_by(state, appliance, refClass) |> 
        mutate(n_distinct_family = n_distinct(family),
               n_distinct_state_avg = n_distinct(state_avg), 
               distinct_family = paste(unique(family), collapse = ","),
               distinct_state_avg = paste(unique(state_avg), collapse = ",")) |>
        filter(n_distinct_family > 1 | n_distinct_state_avg > 1) |> 
        group_by(state, appliance, refClass, family, state_avg) |> 
        mutate(state_avg_count = n_distinct(state_avg),
               family_count = n()) |> 
        ungroup() |> 
        group_by(state, appliance, refClass, state_avg) |> 
        mutate(state_avg_count = n()) |> 
        distinct(state, appliance, refClass, family, state_avg, family_count, state_avg_count) |> 
      arrange(state,appliance,refClass,-family_count) |> 
      rowwise() |> 
      mutate(fam_in = family %in% correct_values$family,
             state_in = state_avg %in% correct_values$state_avg) |>
       kable() # print(n=25)

# Function to check values
check_values <- function(data, reference) {
  # Join the datasets
  data %>%
    left_join(reference |> rename(family_true=family,state_true=state_avg), by = c("state" = "state_abrv", "appliance" = "appliance")) %>%
    mutate(
      # Convert to numeric to ensure proper comparison
      family_numeric = as.numeric(family),
      state_avg_numeric = as.numeric(state_avg),
      # Flag discrepancies (using a small tolerance for floating point differences)
      family_discrepant =family_numeric != family_true,
      state_avg_discrepant = state_avg_numeric != state_true
    )
}

discrepant_counts <- s1 %>%
  # Remove total kWh as specified
  #filter(appliance != "Total kWh") %>%
  # Use only one plan to avoid duplicates
  filter(plan == "plan1") %>%
  # Check against correct values
  check_values(correct_values) %>%
  select(id,id2,starts_with("sta"),starts_with("fam"),appliance,refClass) |> 
  # Filter to show only discrepant cases
  filter(family_discrepant | state_avg_discrepant) %>%
  # Get unique subject IDs and their discrepancies
  group_by(refClass,state) %>%
  summarise(
    discrepant_states = paste(unique(state), collapse = ", "),
    discrepant_appliances = paste(unique(appliance), collapse = ", "),
    n_family_discrepancies = sum(family_discrepant),
    n_state_avg_discrepancies = sum(state_avg_discrepant),
    .groups = 'drop'
  )

# Identify discrepant subjects
discrepant_subjects <- s1 %>%
  # Remove total kWh as specified
  #filter(appliance != "Total kWh") %>%
  # Use only one plan to avoid duplicates
  filter(plan == "plan1") %>%
  # Check against correct values
  check_values(correct_values) %>%
  select(id,id2,refClass,starts_with("sta"),starts_with("fam"),appliance) |> 
  # Filter to show only discrepant cases
  filter(family_discrepant | state_avg_discrepant) %>%
  # Get unique subject IDs and their discrepancies
  group_by(id, id2,refClass) %>%
  summarise(
    discrepant_states = paste(unique(state), collapse = ", "),
    discrepant_appliances = paste(unique(appliance), collapse = ", "),
    n_family_discrepancies = sum(family_discrepant),
    n_state_avg_discrepancies = sum(state_avg_discrepant),
    .groups = 'drop'
  )

discrepant_counts |> gt()
discrepant_subjects |> gt()

s1_discrep_ids <- discrepant_subjects |> pull(id) |> unique()

s1_discrep_ids

# for every combination of state, appliance, and refClass, calculate the number of distinct values of family, and of state_avg
s1 |> group_by(state,appliance,refClass) |> 
  filter(appliance != "Total kWh") %>%
  filter(!(id %in% s1_discrep_ids)) |>
  mutate(state_avg=as.numeric(state_avg)) |> 
  summarise(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  filter(n_distinct_family>1 | n_distinct_state_avg>1)

# <!-- issue in slides where colorado cooling is shown as 489 in a few cases instead of 498 - all of the data has 498. 498 corresponds with total value as well  -->
s1 |> filter(state=="COL",appliance=="Cooling") |> pull(state_avg) |> unique() #498
s2_long |> filter(state=="Colorado",appliance=="AC") |> pull(state_avg) |> unique() #498

s1 |> select(id,id2,starts_with("sta"),starts_with("fam"),appliance,refClass,plan,value) |> check_values(correct_values) %>% 
filter(plan=="plan1") |> 
select(-state_f_dif,-state_p_dif,-plan,-state.y,-state_avg) |> ungroup() |> as_tibble() |> 
relocate(id,state, appliance,value,family_numeric,family_true,state_avg_numeric,state_true, family_discrepant,state_avg_discrepant) |>
# mutate to turn applilance, state and id columns into factors
mutate(across(c(appliance,state,id),factor)) |>
#reactable::reactable()
  DT::datatable(extensions = 'Buttons',
    options = list(
       dom = 'Blfrtip',
       buttons = list('copy', 'csv', 'excel', 'pdf', 'print'),
      pageLength = 25,
      autoWidth = TRUE
    ),
    filter = 'top')

```

```{r}

haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |> select(P3TEX15) |> unique()

s1 |> filter(id==18,plan=="plan1")

haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |>
  filter(ResponseId == "R_vB56XspT1LqhrkB") |>
  # Select columns with non-NA values and not empty strings, handling character and numeric columns separately
  select(
    where(~ is.numeric(.) && !all(is.na(.))),
    where(~ is.character(.) && !all(is.na(.)) && !all(. == ""))
  ) |>
  # Remove columns that start with "AIN"
  select(-starts_with("AIN"))

haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))  |> select_if(~any(. ==5124)) |> 
  # filter out any NAs from any column that is left (we don't know the column names)
  filter_all(any_vars(!is.na(.)))

```

```{r}

s2_long %>%
  filter(appliance != "TOTAL", plan == "plan1") %>%
  group_by(appliance, rounded, state) %>%
  mutate(
    distinct_families = n_distinct(family),
    color_flag = distinct_families > 1  # Flag as TRUE if more than 1 distinct value
  ) %>%
  ungroup() %>%
  # Continue grouping for plotting
  group_by(appliance, family, rounded, state,color_flag) %>%
  summarise(n = n(), .groups = "drop")

```