### 2024-11-20 mpa/mpa_analysis.qmd

---
title: "MPA abstract analysis"
execute:
    warning: false
format:
    html: 
        grid:
            sidebar-width: 220px
            body-width: 1200px
            margin-width: 170px
            gutter-width: 1.0rem
    hugo-md:
        include: true
        html-math-method: mathjax
        output-file: mpa_hugo.md
    gfm:
        echo: true
        output-file: mpa_gfm.md
---

```{r}
pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())
options(brms.backend="cmdstanr",mc.cores=4)

s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
```

## Fit Ordinal model accuracy_level ~ refClass + (1|id) + (1|state)
```{r}

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass + (1|id) + (1|state),
    data = s1_agg,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    control = list(adapt_delta = 0.99), 
    prior = c(prior(normal(0, 2), class = "Intercept"), 
                prior(normal(0, 1.5), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
)

summary(ordinal_model_s1)

pp_check(ordinal_model_s1)

describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
    filter(stringr::str_detect(Parameter, "b_")) |> 
    mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
    mutate(across(where(is.numeric), round, 3)) |> 
    kable(booktabs = TRUE)

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, .category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

# Convert log-odds to odds ratios
posterior_samples <- as.data.frame(ordinal_model_s1)
odds_ratios <- data.frame(
  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
)

# Calculate summary statistics
odds_ratio_summary <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh"),
    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
                    mean(odds_ratios$USD_vs_kWh)),
    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
                quantile(odds_ratios$USD_vs_kWh, 0.025)),
    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
                quantile(odds_ratios$USD_vs_kWh, 0.975))
)

odds_ratio_summary |> kable()
# |comparison        | odds_ratio| ci_lower| ci_upper|
# |:-----------------|----------:|--------:|--------:|
# |Percentage vs kWh |        2.3|     0.62|      6.2|
# |USD vs kWh        |       14.1|     4.08|     38.0|

odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")

# Plot predicted probabilities
ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = .category)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
                    position = position_dodge(width = 0.9), width = 0.2) +
    labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
    ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
    theme_minimal()

```

|comparison        | odds_ratio| ci_lower| ci_upper|
|:-----------------|----------:|--------:|--------:|
|Percentage vs kWh |        2.3|     0.62|      6.2|
|USD vs kWh        |       14.1|     4.08|     38.0|

# Study 2

```{r}

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, state,refClass,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

observed_props2 <- s2_agg1 |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  pivot_wider(
    names_from = accuracy_level,
    values_from = c(n, prop)
  )

  # Calculate proportions
prop_acc_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

# Plot
ggplot(prop_acc_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(title = "Distribution of Accuracy Levels by Reference Class",
        x = "Accuracy Level",
        y = "Probability",
        fill = "Reference Class") +
    theme_minimal() +
    scale_fill_brewer(palette = "Set1")

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal() + facet_wrap(~rounded+pct_goal)

prop4_acc_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level,rounded,pct_goal) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

# Plot

s2_agg1 %>%
    group_by( accuracy_level,rounded) %>%
    summarise(count = n()) %>%
    group_by(rounded) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

ggplot(prop4_acc_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(title = "Distribution of Accuracy Levels by Reference Class",
        x = "Accuracy Level",
        y = "Probability",
        fill = "Reference Class") +
    theme_minimal() +
    scale_fill_brewer(palette = "Set1") + facet_wrap(~rounded+pct_goal)

ggplot(prop4_acc_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(title = "Distribution of Accuracy Levels by Reference Class",
        x = "Accuracy Level",
        y = "Probability",
        fill = "Reference Class") +
    theme_minimal() +
    scale_fill_brewer(palette = "Set1")

```

# no interaction model used in MPA
```{r}

s2_agg1 <- s2_agg1 %>%
  mutate(
    refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")),
    rounded = factor(rounded, levels = c("Not Rounded", "Rounded")),
    pct_goal = factor(pct_goal, levels = c("10%", "15%"))
  )

ordinal_model_s2_probit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("probit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_op.rds')) # Cache for efficiency
)

summary(ordinal_model_s2_probit)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept[1]          -1.12      0.38    -1.88    -0.37 1.01      686     1070
# Intercept[2]          -0.28      0.38    -1.04     0.47 1.01      693     1085
# Intercept[3]           1.81      0.38     1.04     2.56 1.01      702     1248
# refClassPercentage     0.58      0.45    -0.32     1.49 1.01      446      741
# refClassUSD            1.19      0.47     0.29     2.11 1.01      583     1087
# roundedRounded        -0.37      0.10    -0.57    -0.17 1.00     5613     2928
# pct_goal15%           -0.26      0.10    -0.46    -0.06 1.00     5136     2842

pred_summary_s2 <- ordinal_model_s2_probit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

pred_summary_s2 |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class")

# Extract posterior samples
posterior_samples_s2 <- as.data.frame(ordinal_model_s2_probit)

odds_ratios_s2 <- posterior_samples_s2 %>%
    transmute(
        refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
        refClass_USD_vs_kWh = exp(b_refClassUSD),
        rounded_Yes_vs_No = exp(b_roundedRounded),
        pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
    )

# Calculate summary statistics
odds_ratio_summary_s2 <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
    odds_ratio = c(
        mean(odds_ratios_s2$refClass_Percentage_vs_kWh),
        mean(odds_ratios_s2$refClass_USD_vs_kWh),
        mean(odds_ratios_s2$rounded_Yes_vs_No),
        mean(odds_ratios_s2$pct_goal_15_vs_10)
    ),
    ci_lower = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.025),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.025),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.025),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.025)
    ),
    ci_upper = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.975),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.975),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.975),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.975)
    )
)

#             comparison odds_ratio ci_lower ci_upper
# 1    Percentage vs kWh       1.99     0.73     4.43
# 2           USD vs kWh       3.69     1.33     8.29
# 3       Rounded vs Not       0.70     0.57     0.84
# 4 15% Goal vs 10% Goal       0.78     0.63     0.94

odds_ratio_summary_s2 |> kable()

odds_ratio_summary_s2 |> pander::pandoc.table(caption="Study 2: Odds ratios of accuracy")

ggplot(pred_summary, 
        aes(x = .category, y = mean_prob, fill = refClass )) +
    geom_col(position = position_dodge(width = 0.8)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
                    position = position_dodge(width = 0.8)) +
    labs(x = "Reference Class", 
        y = "Predicted Probability",
        fill = "Accuracy Level",
        title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
    scale_fill_brewer(palette = "RdYlBu") +
    theme_minimal()

```

|comparison           | odds_ratio| ci_lower| ci_upper|
|:--------------------|----------:|--------:|--------:|
|Percentage vs kWh    |       1.99|     0.73|     4.43|
|USD vs kWh           |       3.69|     1.33|     8.29|
|Rounded vs Not       |       0.70|     0.57|     0.84|
|15% Goal vs 10% Goal |       0.78|     0.63|     0.94|