### 2024-11-20 paper.qmd

---
title: Representation Project
subtitle: "Planning to Save Energy: How Information Format Affects Accuracy"
toc: false
sidebar: false
repo-actions: false
---

```{=html}
<p class="text-center">
  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="manuscript">
  <i class="fa-solid fa-file-lines"></i> &ensp;HTML
  </a>

  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="_blank">
  <i class="fa-solid fa-arrow-up-right-from-square"></i>&ensp;HTML <small>(new window)</small>
  </a>

  <a class="btn btn-danger btn-sm mb-1" href="/manuscript/output/rep_study.pdf" target="manuscript">
  <i class="fa-solid fa-file-pdf"></i>&ensp;PDF
  </a>

  <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/rep_study.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
  </a>

</p>

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:88%;width:100%;border:2px solid #ccc;" height="88%" width="100%" name="manuscript"></iframe>
```