### 2024-12-08 docs/misc/log_error.qmd
big update to results section. Added a few paragraphs to Intro.

```{r}

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

```

```{r}

pacman::p_load(lmerTest) #provides p-values in lmer summary
# Log transformation of absolute error (add a small constant to avoid log(0))
s1_agg$log_abs_error <- log(s1_agg$abs_error + 0.001)

s1_agg$log_abs_error <- log(s1_agg$abs_error + 0.001)
# Linear mixed-effects model with log-transformed error
error_model <- lmer(log_abs_error ~ refClass + (1|id) + (1|state), data = s1_agg)
summary(error_model)

error_model_table <- broom.mixed::tidy(error_model) %>%
mutate(Estimate = exp(estimate)) %>%
rename('Coefficient' = term,
        'Estimate (Exp)' = Estimate) %>%
select(effect, Coefficient, `Estimate (Exp)`, std.error, statistic)

error_model_table %>%
kable(escape = FALSE, booktabs = TRUE, align = c("lccc"), caption = "Effect of Reference Class on Log-Transformed Error Magnitude (Mixed Model)") %>%
kable_styling(full_width = FALSE)

# diagnostic plots to assess model assumptions
plot(error_model) # residual plot
qqnorm(resid(error_model))
qqline(resid(error_model))

#rename coefficients and add stars to indicate significance
coef_map2 <- c("refClassPercentage" = "Reference Class: Percentage",
              "refClassUSD" = "Reference Class: USD",
              "(Intercept)" = "Intercept")
error_model_table <- modelsummary(error_model,
                                  stars=c('*' = .1, '**' = .05, '***' = .01),
                                  coef_rename = coef_map2,
                                  gof_omit = 'AIC|BIC|Log.Lik.|RMSE|Std.Dev.|VC|sigma|R2 Within',
                                  title = 'Linear Mixed-Effects Model on Log-Transformed Error')

error_model_table
```

Table: Effect of Reference Class on Log-Transformed Error Magnitude (Mixed Model)

|effect   |    Coefficient     | Estimate (Exp) | std.error |statistic |
|:--------|:------------------:|:--------------:|:---------:|:---------|
|fixed    |    (Intercept)     |      0.01      |   0.22    |-19.1     |
|fixed    | refClassPercentage |      1.64      |   0.32    |1.6       |
|fixed    |    refClassUSD     |      4.12      |   0.29    |4.8       |
|ran_pars |  sd__(Intercept)   |      6.26      |    NA     |NA        |
|ran_pars |  sd__(Intercept)   |      1.12      |    NA     |NA        |
|ran_pars |  sd__Observation   |      2.73      |    NA     |NA        |

Fixed effects:
                   Estimate Std. Error      df t value             Pr(>|t|)    
(Intercept)          -4.244      0.222 162.883  -19.11 < 0.0000000000000002 ***
refClassPercentage    0.496      0.316 232.000    1.57                 0.12    
refClassUSD           1.416      0.295 231.832    4.80            0.0000028 ***

```{r}

error_model2 <- glmer(
  log_abs_error ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = gaussian(link = "identity")
)
summary(error_model2)

# Publication-worthy table
error_model_table <- broom.mixed::tidy(error_model2) %>%
  mutate(Estimate = exp(estimate)) %>%
  rename('Coefficient' = term,
         'Estimate (Exp)' = Estimate) %>%
  select(effect,Coefficient, `Estimate (Exp)`,std.error,statistic)

error_model_table %>%
  kable(escape = FALSE, booktabs = TRUE, align = c("lccc"), caption = "Effect of Reference Class on Log-Transformed Error Magnitude (Mixed Model)")

```

Table: Effect of Reference Class on Log-Transformed Error Magnitude (Mixed Model)

|effect   |    Coefficient     | Estimate (Exp) | std.error |statistic |
|:--------|:------------------:|:--------------:|:---------:|:---------|
|fixed    |    (Intercept)     |      0.01      |   0.22    |-19.1     |
|fixed    | refClassPercentage |      1.64      |   0.32    |1.6       |
|fixed    |    refClassUSD     |      4.12      |   0.29    |4.8       |
|ran_pars |  sd__(Intercept)   |      6.26      |    NA     |NA        |
|ran_pars |  sd__(Intercept)   |      1.12      |    NA     |NA        |
|ran_pars |  sd__Observation   |      2.73      |    NA     |NA        |

## Bayesian version

```{r}

log_error_model_s1 <- brm(
  log_abs_error ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s1.rds')) 
)

summary(log_error_model_s1)

# exp and make table
log_error_model_table <- as.data.frame(describe_posterior(log_error_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
  setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
  mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
    mutate(across(where(is.numeric), exp)) |>
  kable(booktabs = TRUE)

# plot density of raw data
s1_agg |> 
  ggplot(aes(x = abs_error)) +
  geom_density(fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ refClass, scales = "free") +ggtitle("Error density by reference class")

# plot density of log-transformed data
s1_agg |> 
  ggplot(aes(x = log_abs_error)) +
  geom_density(fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ refClass, scales = "free") + ggtitle("Log-transformed error density by reference class")

# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -4.23      0.25    -4.72    -3.73 1.01      630      722
# refClassPercentage     0.51      0.31    -0.08     1.12 1.00      583     1050
# refClassUSD            1.42      0.30     0.86     2.01 1.01      502      795

pp_check(log_error_model_s1, type = "dens_overlay_grouped", group="refClass", fatten = 2,ndraws=100) +
  ggtitle("Posterior Predictive Check - log_error_model_s1") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))

```

|Parameter          | Estimate| CI_Lower| CI_Upper|  pd|
|:------------------|--------:|--------:|--------:|---:|
|Intercept          |     0.01|     0.01|     0.02| 2.7|
|refClassPercentage |     1.66|     0.92|     3.08| 2.6|
|refClassUSD        |     4.15|     2.36|     7.49| 2.7|

```{r}

model_list <- list("Ordinal Model (Accuracy)" = ordinal_model_s1,
                   "Gamma Model (Error Magnitude)" = log_error_model_s1)

ms_table <- modelsummary(model_list,
                         exponentiate = c(TRUE, FALSE), # first model OR, second model no transform
                         statistic = "{estimate} [{conf.low}, {conf.high}]",
                         output = "kableExtra",
                         booktabs = TRUE,
                         align = "lcc",
                         title = "Model Comparison: Accuracy (Ordinal) and Error Magnitude (Gamma)",
                         stars = TRUE) %>%
  kable_styling(full_width = FALSE)

print(ms_table)

```

## quantile regression

```{r}

library(quantreg)

# Fit quantile regression for median and upper quantiles
error_qr <- rq(abs_error ~ refClass, data = s1_agg, tau = c(0.5, 0.9))

# Summarize results
summary(error_qr)

# Publication-worthy table

```

Call: rq(formula = abs_error ~ refClass, tau = c(0.5, 0.9), data = s1_agg)

tau: [1] 0.9

Coefficients:
                   coefficients lower bd upper bd
(Intercept)         0.3720       0.2818   0.5205 
refClassPercentage -0.1180      -0.2234  -0.0031 
refClassUSD        -0.0140      -0.1405   0.1016

```{r}

#--------------------------------------
# 2. Model the effect of refClass on error magnitude
#--------------------------------------
# Considering abs_error as a continuous variable:
# Because this variable might be zero or very small, check its range. If strictly positive, Gamma can be used.
# If zero values exist, consider adding a small constant or using a zero-inflated model.
# Here, we assume abs_error > 0.

# Fit a Gamma regression model with a log link:
error_model_gamma <- brm(
  abs_error ~ refClass + (1 | id) + (1 | state),
  data = s1_agg,
  family = Gamma(link = "log"),
  cores = 4, iter = 5000, control = list(adapt_delta = 0.95),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 3), class = "Intercept")
  )
)
```

```{r}

 mix <- mixture(gaussian, gaussian)

    # Priors for the mixture model (adjust as needed)
    prior_mix <- c(
      prior(normal(0, 4), class = "b", dpar = "mu1"),
      prior(normal(0, 4), class = "b", dpar = "mu2"),
      prior(cauchy(0, .5), class = "sd", dpar = "mu1"),
      prior(cauchy(0, .5), class = "sd", dpar = "mu2"),
      prior(dirichlet(2, 2), class = "theta") # Prior for mixing proportions
    )

# Fit the mixture model
mixture_model_s1 <- brm(
  log_abs_error ~ refClass + (1|id),
  data = s1_agg,
  family = mix,
  prior = prior_mix,
  iter = 2000, chains = 2, cores = 2,
  file = paste0(here::here("data/model_cache",'mixture_model_s1.rds'))
)

summary(mixture_model_s1)
pp_check(mixture_model_s1, type = "dens_overlay_grouped", group="refClass", ndraws=100) + ggtitle("mixture_model_s1") + theme_minimal()

pp_check(mixture_model_s1, type = "dens_overlay_grouped", group="refClass", ndraws=100) 
```

# S2

```{r}

log_error_model_s2_c1 <- brm(
  log_abs_error ~ refClass*calc + (1|id) + (1|state),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1.rds')) 
)

summary(log_error_model_s2_c1)
conditional_effects(log_error_model_s2_c1)

log_error_model_s2_c1rf <- brm(
  log_abs_error ~ refClass + (1|id) + (1|state) + (1|calc),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf.rds')) 
)

summary(log_error_model_s2_c1rf)

log_error_model_s2_cAdd_int3 <- brm(
  log_abs_error ~ (refClass*rounded*pct_goal)+calc + (1|id) + (1|state),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_cAdd_int3.rds')) 
)

summary(log_error_model_s2_cAdd_int3)
conditional_effects(log_error_model_s2_cAdd_int3)

log_error_model_s2_c1rf_int3 <- brm(
  log_abs_error ~ refClass*rounded*pct_goal + (1|id) + (1|state) + (1|calc),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf_int3.rds')) 
)

summary(log_error_model_s2_c1rf_int3)

conditional_effects(log_error_model_s2_c1rf_int3)

log_error_model_s2_c1rf_add3 <- brm(
  log_abs_error ~ refClass+rounded+pct_goal + (1|id) + (1|state) + (1|calc),
  data = s2_agg1,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf_add3.rds')) 
)

summary(log_error_model_s2_c1rf_add3)
conditional_effects(log_error_model_s2_c1rf_add3)

```

```{r}

```