### 2024-12-08 manuscript/appendix.qmd

```{r}
#| cache: false
pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
    ggplot2,gt,knitr,kableExtra,ggh4x,patchwork, ggridges,ggstance,lme4,flextable,pander)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))

theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

```

## Study 1

### Amount of change

```{r}

# plot mean_pct_change distribution
s1_agg4 |> ggplot(aes(x = refClass, y = mean_pct_change, fill = refClass)) +
    geom_boxplot() +
    labs(title = "S1. Mean % Change by Reference Class",
        x = "Reference Class",
        y = "Mean % Change",
        fill = "Reference Class") +
    theme_minimal()

# plot mean_pct_change density
s1_agg4 |> ggplot(aes(x = mean_pct_change, fill = refClass)) +
    geom_density(alpha = 0.5) +
    labs(title = "S1. Mean % Change Density by Reference Class",
        x = "Mean % Change",
        fill = "Reference Class") +
    theme_minimal()

```

## Study2

### Goal Manipulation

```{r}

ggplot(s2_agg1, aes(y = pct_goal, x = log_abs_error, color = calc)) +
  geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
                      position = position_raincloud(width = 0.05, height = 0.1,
                                                    ygap = 0.05)) +
  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
 # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
  guides(fill = "none") +
  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "") +
  theme(legend.position = "top")

ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = pct_goal)) +
    geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
    labs(title = "S1. % of Participants within each Accuracy Bin",
        x = "Accuracy Level",
        y = "Percentage of Participants",
        fill = "Reference Class") +
    theme_minimal()

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = pct_goal)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = pct_goal)) +
geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
scale_y_continuous(labels = scales::percent) +
facet_wrap(~refClass) +
labs(title = "Accuracy Levels by Reference Class",
      x = "Accuracy Level",
      y = "Percentage of Participants",
      fill = "Reference Class") +
theme_minimal()

```

### Rounding

```{r}

ggplot(s2_agg1, aes(y = refClass, x = log_abs_error, color = rounded)) +
  geom_density_ridges(aes(fill = rounded), alpha = 0.2, scale = 0.5,
                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
                      position = position_raincloud(width = 0.05, height = 0.1,
                                                    ygap = 0.05)) +
  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
 # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
  guides(fill = "none") +
  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "") +
  theme(legend.position = "top")

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = rounded)) +
geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
scale_y_continuous(labels = scales::percent) +
facet_wrap(~refClass) +
labs(title = "Accuracy Levels by Reference Class",
      x = "Accuracy Level",
      y = "Percentage of Participants",
      fill = "Reference Class") +
theme_minimal()

```

### Tables

### Distance plots
```{r}
#| label: tbl-s2-agg
#| tbl-cap: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy))

s2_agg4 %>% ggplot(aes(x=refClass,y=mean_log_abs_error,fill=refClass)) + 
  stat_summary(fun=mean, geom="bar") +
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
  geom_jitter(alpha = 0.2, width = 0.2, height = 0)

s2_agg4 |> 
  ggplot(aes(x = mean_log_abs_error,fill = refClass)) +
  geom_density(alpha = 0.7)

```