### 2024-12-08 misc/odds_ratio.qmd

```{r}

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

```

```{r}
#| label: tbl-s1-ord
#| tbl-cap: "Study 1: Odds ratios of for group comparisons. "

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass + (1|id) + (1|state),
    data = s1_agg,
    family = cumulative("logit"),
    cores = 4,
    iter = 5000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
)
# 
# ordinal_model_s1_null <- brm(
#     accuracy_level ~  (1|id) + (1|state),
#     data = s1_agg,
#     family = cumulative("logit"),
#     cores = 4,
#     iter = 5000,
#     save_pars = save_pars(all = TRUE),
#     control = list(adapt_delta = 0.99), 
#     prior = c(prior(normal(0, 3), class = "Intercept")),
#     file = paste0(here::here("data/model_cache",'s1_null.rds')) 
# )

#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
  setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
  mutate(Parameter = stringr::str_remove(Parameter, "b_"))

# ordinal_model_s1 |> emmeans(~refClass) |> contrast(method="pairwise") |>  
#   gather_emmeans_draws() %>%
#   mean_hdi(.width = .95)

# describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
#     filter(stringr::str_detect(Parameter, "b_")) |> 
#     mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
#     mutate(across(where(is.numeric), round, 3)) |> 
#     kable(booktabs = TRUE)

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, Category=.category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

# Convert log-odds to odds ratios
posterior_samples <- as.data.frame(ordinal_model_s1)
odds_ratios <- data.frame(
  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
)

# Calculate summary statistics
odds_ratio_summary <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh"),
    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
                    mean(odds_ratios$USD_vs_kWh)),
    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
                quantile(odds_ratios$USD_vs_kWh, 0.025)),
    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
                quantile(odds_ratios$USD_vs_kWh, 0.975))
)

odds_ratio_summary |> kable(escape=FALSE,booktabs=TRUE,align=c("l")) 
# |comparison        | odds_ratio| ci_lower| ci_upper|
# |:-----------------|----------:|--------:|--------:|
# |Percentage vs kWh |        2.3|     0.62|      6.2|
# |USD vs kWh        |       14.1|     4.08|     38.0|

#odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")

# odds ratios of fixed effects
as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
    rownames_to_column(var = "Parameter") %>%
    mutate(across(where(is.numeric), exp)) |>
    filter(!stringr::str_detect(Parameter, "Intercept"))

describe_posterior(ordinal_model_s1, centrality = "Median")[, c(1,2,4,5)] %>%
  as_tibble() |> # remove rows with Intercept
  mutate(across(where(is.numeric), exp)) |>
  filter(!stringr::str_detect(Parameter, "Intercept")) |>
  mutate(Parameter = stringr::str_remove(Parameter, "b_"))

or_table <- data.frame(exp(fixef(ordinal_model_s1)))
or_table |>
  rownames_to_column(var = "Term") |>
  as_flextable() |>
  set_header_labels(values = list(
    OR = "Odds Ratio",
    `2.5 %` = "Lower 95% CI",
    `97.5 %` = "Upper 95% CI"
  )) |>
  flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
  align(align = 'center', part = 'all') |>
  fontsize(size = 10, part = "all")

posterior_draws <- as_draws_df(ordinal_model_s1) |> select(starts_with("b_"))

# Compute posterior summaries and exponentiate to get odds ratios
or_summary <- posterior_draws %>%
  summarise(across(starts_with("b_"),
                   list(mean = ~mean(.x),
                        sd = ~sd(.x),
                        lower95 = ~quantile(.x, 0.025),
                        upper95 = ~quantile(.x, 0.975)))) %>%
  tidyr::pivot_longer(cols = everything(),
                      names_to = c("parameter", ".value"),
                      names_pattern = "(b_.*)_(.*)") %>%
  mutate(odds_ratio = exp(mean),
         lower95_or = exp(lower95),
         upper95_or = exp(upper95))

or_table <- or_summary %>%
  select(parameter, odds_ratio, lower95_or, upper95_or) %>%
  mutate(parameter = gsub("b_", "", parameter)) %>%
  rename("Predictor" = parameter,
         "OR" = odds_ratio,
         "2.5% CI" = lower95_or,
         "97.5% CI" = upper95_or) %>%
  kbl(digits = 2, booktabs = TRUE, align = "lccc",
      caption = "Odds Ratios for Ordinal Model") %>%
  kableExtra::kable_styling(full_width = FALSE)

print(or_table)

```

```{r}

summary(ordinal_model_s1)

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check - ordinal_model_s1") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

pp_check(ordinal_model_s1, type = "dens_overlay_grouped", group="refClass", fatten = 2,ndraws=100) +   scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check - ordinal_model_s1") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}

multinomial_model_s1 <- brm(
  accuracy_level ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = categorical(), # Note: categorical() for multinomial
  cores = 4,
  iter = 1000,
  save_pars = save_pars(all = TRUE),
  control = list(adapt_delta = 0.99, max_treedepth=12),
#   prior = c(prior(normal(0, 3), class = "Intercept"), 
#             prior(normal(0, 3), class = "b")),
  file = paste0(here::here("data/model_cache",'s1_multinomial.rds')) 
)

summary(multinomial_model_s1)
pp_check(multinomial_model_s1, type = "bars_grouped", group="refClass", fatten = 2,ndraws=100) + labs(title = "Study 1: Posterior predictive check for the multinomial model") + theme_minimal()

```

```{r}
#| eval: false

loo_ordinal <- LOO(ordinal_model_s1, moment_match=TRUE)
loo_multinomial <- LOO(multinomial_model_s1, moment_match=TRUE, reloo=TRUE)
loo_compare(loo_ordinal, loo_multinomial)

loo_ordinal <- LOO(ordinal_model_s1)
loo_multinomial <- LOO(multinomial_model_s1)
loo_compare(loo_ordinal, loo_multinomial)

```

## S1 exact match

```{r}

# Assuming s1_agg is your dataframe from Experiment 1, and it's already loaded

# Fit a Bayesian logistic regression model
exact_match_model <- brm(
  matched_goal ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = bernoulli(link = "logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache", 's1_exact_match_model.rds'))
)
summary(exact_match_model)

exact_match_model_c_add <- brm(
  matched_goal ~ refClass + calc+ (1|id) + (1|state),
  data = s1_agg,
  family = bernoulli(link = "logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_add.rds'))
)
summary(exact_match_model_c_add)

exact_match_model_c_int <- brm(
  matched_goal ~ refClass * calc+ (1|id) + (1|state),
  data = s1_agg,
  family = bernoulli(link = "logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_int.rds'))
)
summary(exact_match_model_c_int )
conditional_effects(exact_match_model_c_int)

```

## outliers

```{r}

identify_outliers_group <- function(x) {
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR_value <- Q3 - Q1
  lower_bound <- Q1 - 3.3 * IQR_value
  upper_bound <- Q3 + 3.3 * IQR_value
  return(x < lower_bound | x > upper_bound)
}

# 2. Apply the Outlier Detection to abs_error and log_abs_error within each group
s1_agg <- s1_agg %>%
  group_by(refClass, calc) %>%
  mutate(
    outlier_abs_error = identify_outliers_group(abs_error),
    outlier_log_abs_error = identify_outliers_group(log_abs_error)
  ) %>%
  ungroup()

#s1_agg |> relocate(outlier_abs_error,.after=abs_error) |> sort(desc(outlier_abs_error)) |> head(5)
s1_agg |> filter(outlier_abs_error) |> select(id,refClass,calc,abs_error,log_abs_error) |> # keep unique id and refClass pairs
  distinct(id,refClass)

# 3. Summarize the Number and Proportion of Outliers by Group
outlier_summary_by_group <- s1_agg %>%
  group_by(refClass, calc) %>%
  summarise(
    total_observations = n(),
    outliers_abs_error = sum(outlier_abs_error),
    percent_outliers_abs_error = round((outliers_abs_error / total_observations) * 100, 2),
    outliers_log_abs_error = sum(outlier_log_abs_error),
    percent_outliers_log_abs_error = round((outliers_log_abs_error / total_observations) * 100, 2)
  ) %>%
  ungroup()

# Display the outlier summary by group
print("Outlier Summary by Reference Class and Calculator Usage:")
print(outlier_summary_by_group)

s1_agg_no_outliers <- s1_agg %>%
  filter(!outlier_abs_error & !outlier_log_abs_error)

# Re-run the exact match comparison without outliers

# Create a contingency table of reference class vs. exact match without outliers
exact_match_table_no_outliers <- s1_agg_no_outliers %>%
  mutate(exact_match = ifelse(matched_goal, "Exact Match", "No Exact Match")) %>%
  count(refClass, exact_match) %>%
  pivot_wider(names_from = exact_match, values_from = n, values_fill = 0)

# Perform Chi-Square Test of Independence without outliers
chi_square_test_no_outliers <- chisq.test(exact_match_table_no_outliers[, -1])

# Display results
print("Chi-Square Test Results (Without Outliers):")
print(chi_square_test_no_outliers)

# Optionally, re-run logistic regression without outliers
logistic_model_no_outliers <- glm(matched_goal ~ refClass, data = s1_agg_no_outliers, family = binomial)
summary(logistic_model_no_outliers)

# Calculate Odds Ratios
odds_ratios_no_outliers <- exp(coef(logistic_model_no_outliers))
print("Odds Ratios (Without Outliers):")
print(odds_ratios_no_outliers)

# Calculate 95% Confidence Intervals for Odds Ratios without outliers
conf_intervals_no_outliers <- exp(confint(logistic_model_no_outliers))
print("95% Confidence Intervals for Odds Ratios (Without Outliers):")
print(conf_intervals_no_outliers)

s1_clean <- s1_agg %>%
  # Remove trials that are outliers in abs_error or log_abs_error
  #filter(!(outlier_abs_error | outlier_log_abs_error))
  # remove any id that has any outlier_abs_error
  filter(!(id %in% (s1_agg |> filter(outlier_abs_error) |> select(id) |> distinct() |> pull(id))))

length(unique(s1_agg$id)) 
length(unique(s1_clean$id))

# 2. Summary of Cleaned Data
cleaned_summary <- s1_clean %>%
  group_by(refClass, calc) %>%
  summarise(
    total_observations = n(),
    exact_matches = sum(matched_goal),
    percent_exact = round((exact_matches / total_observations) * 100, 2),
    mean_abs_error = mean(abs_error, na.rm = TRUE),
    mean_log_abs_error = mean(log_abs_error, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the cleaned data summary
print("Cleaned Data Summary by Reference Class and Calculator Usage:")
print(cleaned_summary)

ordinal_model_s1_clean_calc <- brm(
    accuracy_level ~ refClass +calc + (1|id) + (1|state),
    data = s1_clean,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
)

summary(ordinal_model_s1_clean_calc)

ordinal_model_s1_clean_calc_int <- brm(
    accuracy_level ~ refClass * calc + (1|id) + (1|state),
    data = s1_clean,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_int.rds')) 
)

summary(ordinal_model_s1_clean_calc_int)
conditional_effects(ordinal_model_s1_clean_calc_int)

# logistic regression on exact match with clean data (frequentist)
logistic_model_clean <- glm(matched_goal ~ refClass, data = s1_clean, family = binomial)
summary(logistic_model_clean)

# logistic regression on exact match with clean data (frequentist) - mixed effects
logistic_model_clean_mixed <- glmer(matched_goal ~ refClass + (1|id) + (1|state), data = s1_clean, family = binomial)
summary(logistic_model_clean_mixed)

# logistic regression on exact match with clean data (frequentist) - mixed effects
logistic_model_clean_mixed <- glmer(matched_goal ~ refClass+calc + (1|id) , data = s1_clean, family = binomial)
summary(logistic_model_clean_mixed)

# logistic regression on exact match with clean data (bayesian)
logistic_model_clean_bayes <- brm(
  matched_goal ~ refClass + (1|id) + (1|state),
  data = s1_clean,
  family = bernoulli(link = "logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.97),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = paste0(here::here("data/model_cache", 's1_exact_match_model_clean.rds'))
)
summary(logistic_model_clean_bayes)
conditional_effects(logistic_model_clean_bayes)

```

## outlier from avg.

```{r}

group_stats <- s1_agg |>
  group_by(refClass, calc) |>
  summarise(
    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate participant-level average abs_error
participant_stats <- s1_agg |>
  group_by(id, refClass, calc) |>
  summarise(
    avg_abs_error = mean(abs_error, na.rm = TRUE),
    .groups = "drop"
  )

# Join group stats with participant stats
participant_stats <- participant_stats |>
  left_join(group_stats, by = c("refClass", "calc"))

# Identify participants to exclude based on the 2.5 SD threshold
participants_to_exclude <- participant_stats |>
  mutate(
    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
    exclude = avg_abs_error > exclusion_threshold
  ) |>
  filter(exclude)

ordinal_model_s1_clean_calc <- brm(
    accuracy_level ~ refClass +calc + (1|id) + (1|state),
    data = s1_clean,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
)

s1_agg_filtered <- s1_agg |>
  filter(!(id %in% participants_to_exclude$id))

length(unique(s1_agg$id)) 
length(unique(s1_agg_filtered$id))

ordinal_model_s1_clean_calc <- brm(
    accuracy_level ~ refClass +calc + (1|id) + (1|state),
    data = s1_agg_filtered,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal_grpclean_calc_2.5.rds')) 
)

summary(ordinal_model_s1_clean_calc)
conditional_effects(ordinal_model_s1_clean_calc)

mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
summary(mixed_model_abs_error)
# with p value
library(lmerTest)
mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
summary(mixed_model_abs_error)

```

### clearn log error
```{r}

s1_clean |> 
  ggplot(aes(x = abs_error)) +
  geom_density(fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ refClass, scales = "free") +ggtitle("Error density by reference class")

# plot density of log-transformed data
s1_clean |> 
  ggplot(aes(x = log_abs_error)) +
  geom_density(fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ refClass, scales = "free") + ggtitle("Log-transformed error density by reference class")

# mixed effects regression on abs_error (frequentist)
mixed_model_abs_error <- lmer(abs_error ~ refClass + (1|id) + (1|state), data = s1_clean)
summary(mixed_model_abs_error)

mixed_model_abs_error <- lmer(abs_error ~ refClass*calc + (1|id) + (1|state), data = s1_clean)
summary(mixed_model_abs_error)

```

# S2

```{r}
#| label: tbl-s2-agg
#| tbl-cap: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# overall pct of subjects who matched their goal
s2_agg4 |> group_by(refClass) |>
    summarise(
    'Avg. % meeting goal' = mean(pct),
    'Avg. Deviation From Goal' = median(mean_pct_change),
    # sd = sd(pct),
    # n = n(),
    #se=sd(pct)/sqrt(n)
) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

```

```{r}

ordinal_model_s2_logit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit)

pred_summary_s2 <- ordinal_model_s2_logit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA,ndraws=200) %>%
  group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

# Create a new data frame with the predictor values
new_data <- expand.grid(
  refClass = c("kWh", "Percentage", "USD"),
  id = unique(s2_agg1$id)[1],      # Take first id
  state = unique(s2_agg1$state)[1], # Take first state,
  rounded = unique(s2_agg1$rounded)[1], # Take first rounded
  pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
)

# Generate predicted probabilities
predicted_probabilities <- predict(ordinal_model_s2_logit, newdata = new_data, 
                                  summary = FALSE, ndraws = 1000)

# Calculate the predicted probability of "Exact match"
predicted_probability_exact_match <- mean(predicted_probabilities <= 1)

print(predicted_probability_exact_match)

pred_summary_s2 |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class",split.table=Inf)

# Extract posterior samples
posterior_samples_s2 <- as.data.frame(ordinal_model_s2_logit)

odds_ratios_s2 <- posterior_samples_s2 %>%
    transmute(
        refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
        refClass_USD_vs_kWh = exp(b_refClassUSD),
        rounded_Yes_vs_No = exp(b_roundedRounded),
        pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
    )

# Calculate summary statistics
odds_ratio_summary_s2 <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
    odds_ratio = c(
        mean(odds_ratios_s2$refClass_Percentage_vs_kWh),
        mean(odds_ratios_s2$refClass_USD_vs_kWh),
        mean(odds_ratios_s2$rounded_Yes_vs_No),
        mean(odds_ratios_s2$pct_goal_15_vs_10)
    ),
    ci_lower = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.025),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.025),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.025),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.025)
    ),
    ci_upper = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.975),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.975),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.975),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.975)
    )
)

#             comparison odds_ratio ci_lower ci_upper
# 1    Percentage vs kWh       1.99     0.73     4.43
# 2           USD vs kWh       3.69     1.33     8.29
# 3       Rounded vs Not       0.70     0.57     0.84
# 4 15% Goal vs 10% Goal       0.78     0.63     0.94

odds_ratio_summary_s2 |> kable()

odds_ratio_summary_s2 |> pander::pandoc.table(caption="Study 2: Odds ratios of accuracy")

# Create a new data frame with the predictor values
# new_data <- expand.grid(
#   refClass = c("kWh", "Percentage", "USD"),
#   id = unique(s2_agg1$id)[1],      # Take first id
#   state = unique(s2_agg1$state)[1], # Take first state,
#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
# )

# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
# or_table |>
#   rownames_to_column(var = "Term") |>
#   filter(!stringr::str_detect(Term, "Intercept")) |>
#   as_flextable() |>
#   set_header_labels(values = list(
#     OR = "Odds Ratio",
#     `2.5 %` = "Lower 95% CI",
#     `97.5 %` = "Upper 95% CI"
#   )) |>
#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
#   align(align = 'center', part = 'all') |>
#   fontsize(size = 10, part = "all")

```

# Interaction s2

```{r}

ordinal_model_s2_logit_int2 <- brm(
  accuracy_level ~ refClass +(rounded*pct_goal) + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_int2.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_int2)

conditional_effects(ordinal_model_s2_logit_int2)

ordinal_model_s2_logit_int3 <- brm(
  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_int3)

conditional_effects(ordinal_model_s2_logit_int3)

ordinal_model_s2_logit_c1rf_int3<- brm(
  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state) + (1|calc),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_c1rf_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_c1rf_int3)

conditional_effects(ordinal_model_s2_logit_c1rf_int3)

ordinal_model_s2_logit_c1add_int3<- brm(
  accuracy_level ~ (refClass * rounded*pct_goal) + calc + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.97), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_c1add_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_c1add_int3)

conditional_effects(ordinal_model_s2_logit_c1add_int3)

```