### 2024-12-11 nb/s1_survey.qmd

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,haven,tibble,
               ggplot2,gt,ggh4x,lme4,flextable,kableExtra,brms,emmeans,tidybayes)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())
```

```{r}

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc,els) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc,els) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

```

```{r}

# Scatter plots with regression lines
ggplot(s1_agg4, aes(x = els, y = mean_abs_error)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Add a linear regression line with standard error
  labs(title = "ELS Score vs. Mean Absolute Error",
       x = "ELS Score",
       y = "Mean Absolute Error") +
  theme_bw()

ggplot(s1_agg4, aes(x = els, y = mean_log_abs_error)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "ELS Score vs. Mean Log Absolute Error",
       x = "ELS Score",
       y = "Mean Log Absolute Error") +
  theme_bw()

ggplot(s1_agg4, aes(x = accuracy_level, y = els, fill = accuracy_level)) +
  geom_boxplot() +
  labs(title = "ELS Score by Accuracy Level",
       x = "Accuracy Level",
       y = "ELS Score") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(s1_agg4, aes(x = calc, y = els, fill = calc)) +
          geom_boxplot() +
          labs(title = "ELS Score by Calculator Use",
               x = "Calculator Use",
               y = "ELS Score") +
          theme_bw()

unique(s1_agg4$accuracy_level)

s1_agg4 |> group_by(accuracy_level) |> summarize(n=n())
s1_agg |> group_by(accuracy_level) |> summarize(n=n())

# Calculate correlations
cor.test(s1_agg4$els, s1_agg4$mean_abs_error)
cor.test(s1_agg4$els, s1_agg4$mean_log_abs_error)

# ANOVA
anova_result <- aov(els ~ accuracy_level, data = s1_agg4)
summary(anova_result)

# Tukey's HSD post-hoc test
TukeyHSD(anova_result)

# ordinal regression
polr_model <- MASS::polr(accuracy_level ~ els + refClass + calc, data = s1_agg4)
summary(polr_model)
emmeans::emmeans(polr_model, pairwise ~ els)

library(ordinal)    
library(parameters)

m_equal <- clm(accuracy_level ~ els, 
    data = s1_agg, 
    link="probit")

parameters::model_parameters(m_equal) |> 
  insight::print_html()
summary(m_equal)
# formula: accuracy_level ~ els
# data:    s1_agg

#  link   threshold nobs logLik   AIC     niter max.grad cond.H 
#  probit flexible  1008 -1213.66 2435.32 8(0)  4.05e-11 6.0e+01

# Coefficients:
#     Estimate Std. Error z value            Pr(>|z|)    
# els   -1.957      0.133   -14.7 <0.0000000000000002 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# Threshold coefficients:
#                               Estimate Std. Error z value
# Exact match|0.01-2% error       -2.052      0.099  -20.72
# 0.01-2% error|2.01-15% error    -1.574      0.093  -16.92
# 2.01-15% error|Over 15% error   -0.350      0.083   -4.21

mean2=coef(m_equal)["els"]
Thresholds <- coef(m_equal)[1:3]

polr_model_ref_interaction <- MASS::polr(accuracy_level ~ els * refClass + calc, data = s1_agg4)
summary(polr_model_ref_interaction)
# Example for refClass = USD, calc = No Calculator
new_data <- expand.grid(els = seq(min(s1_agg4$els), max(s1_agg4$els), length.out = 100),
                        refClass = "USD",
                        calc = "No Calculator")

predicted_probs <- predict(polr_model_ref_interaction, newdata = new_data, type = "probs")

# Combine with new_data for plotting
predicted_probs_df <- cbind(new_data, predicted_probs)

# Reshape to long format for ggplot
predicted_probs_long <- tidyr::pivot_longer(predicted_probs_df, 
                                            cols = `Exact match`:`Over 15% error`,
                                            names_to = "accuracy_level",
                                            values_to = "probability")

# Plot
ggplot(predicted_probs_long, aes(x = els, y = probability, color = accuracy_level)) +
  geom_line(linewidth=1.5) +
  labs(title = "Predicted Probabilities of Accuracy Levels by ELS",
       subtitle = "Reference Class = USD, No Calculator",
       x = "ELS Score",
       y = "Probability") +
  theme_bw()

library(lme4)
mixed_model <- lmer(log_abs_error ~ els + (1 | id), data = s1_agg)
emmeans_model <- emmeans::emmeans(mixed_model, pairwise ~ els)
summary(emmeans_model)

```

```{r}

s1_els_log_error <- brm(
    log_abs_error ~ els + (1|id) + (1|state),
    data = s1_agg,
    family = gaussian(),
    cores = 4,
    iter = 2000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_els_log_error.rds')) 
)

summary(s1_els_log_error)
# Regression Coefficients:
#           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept    -1.65      0.23    -2.12    -1.18 1.00      921     1465
# els          -2.35      0.27    -2.88    -1.81 1.00      753      997

conditional_effects(s1_els_log_error)

# Create the conditional effects plot
conditional_effects_plot <- conditional_effects(s1_els_log_error)

# Extract the data for plotting
plot_data <- conditional_effects_plot[[1]]

# Create the plot
ggplot(plot_data, aes(x = els, y = estimate__)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.2) +
  labs(
    x = "Energy Literacy Score",
    y = "Log Absolute Error",
    title = "Conditional Effect of Energy Literacy on Log Absolute Error"
  ) +
  theme_minimal()

```

# Initial exploration from raw data

## Survey Questions and Energy Literacy Survey

Table 1: Survey Questions and Energy Literacy Survey

| Question | Variable | Values |
|----------|----------|--------|
| "Did you use a calculator?" Used a calculator | MATH01 | 1 = Selected <br> Missing = Not selected |
| Used a paper/pen | MATH02 | 1 = Selected <br> Missing = Not selected |
| Used something else | MATH03 | 1 = Selected <br> Missing = Not selected |
| Used something else text entry | MATH04 | 1 = Selected <br> Missing = Not selected |
| ENERGY LITERACY SURVEY |  |  |
| Dewaters and Powers 2011. Assesses participant knowledge of energy names, use, and sources <br> Question about electrical energy units (kWh) | ELS01 | 1 = "Kilowatt (kW)" <br> 2 = "Kilowatt-hours (kWh)" <br> 3 = "British Thermal Units (BTU)" <br> 4 = "Volts (V)" <br> 5 = "Horsepower (HP)" |
| Question about energy consumed and appliance power rating | ELS02 | 1 = "Multiplied by the cost of electricity" <br> 2 = "Added to the cost of electricity" <br> 3 = "Multiplied by the time it's used" <br> 4 = "Divided by the time it's used" <br> 5 = "Added to the time it's used" |
| Question about incandescent lightbulb conversion | ELS03 | 1 = "Electrical energy to radiant energy (light)" <br> 2 = "Chemical energy to radiant energy (light)" <br> 3 = "Electrical energy to radiant energy (light) and thermal energy (heat)" <br> 4 = "Chemical energy to radiant energy (light) and thermal energy (heat)" <br> 5 = "Electrical energy to radiant energy (light) and mechanical energy" |
| Question about reason to buy energy star appliances | ELS04 | 1 = "ENERGY STAR appliances are usually bigger" <br> 2 = "ENERGY STAR appliances cost more" <br> 3 = "ENERGY STAR appliances use less energy" <br> 4 = "ENERGY STAR appliances are more modern looking" <br> 5 = "ENERGY STAR appliances cost less" |
| Question about which appliances uses the most energy | ELS05 | 1 = "Refrigerating food and beverages" <br> 2 = "Washing and drying clothing" <br> 3 = "Heating and cooling rooms" <br> 4 = "Heating and cooling water" <br> 5 = "Lighting the home" |
| Question about which appliance uses the most electricity | ELS06 | 1 = "Lights" <br> 2 = "Refrigerator" <br> 3 = "Telephone" <br> 4 = "Television" <br> 5 = "Computer" |
| Question about which source provides most electricity in the US | ELS07 | 1 = "Nuclear power" <br> 2 = "Burning petroleum" <br> 3 = "Burning coal" <br> 4 = "Solar energy" <br> 5 = "Water (hydro) power" |
| Question about the problem with electric cars | ELS08 | 1 = "Most electricity is currently produced from fossil fuels (coal, oil, natural gas)" <br> 2 = "Switching to electric cars will make unemployment rates go up" <br> 3 = "It has been proven that it is impossible to build electric cars in great quantities" <br> 4 = "You can't use electricity to operate a car" <br> 5 = "There is nothing wrong with this idea" |

### Correct answers to the questions in the DeWaters & Powers (2011) energy literacy survey:

	1.	ELS01 - Electrical energy units (kWh):
	•	Correct Answer: 2 = Kilowatt-hours (kWh)
	2.	ELS02 - Energy consumed and appliance power rating:
	•	Correct Answer: 3 = Multiplied by the time it’s used
	3.	ELS03 - Incandescent lightbulb conversion:
	•	Correct Answer: 3 = Electrical energy to radiant energy (light) and thermal energy (heat)
	4.	ELS04 - Reason to buy ENERGY STAR appliances:
	•	Correct Answer: 3 = ENERGY STAR appliances use less energy
	5.	ELS05 - Which appliances use the most energy:
	•	Correct Answer: 3 = Heating and cooling rooms
	6.	ELS06 - Which appliance uses the most electricity:
	•	Correct Answer: 2 = Refrigerator
	7.	ELS07 - Which source provides most electricity in the US:
	•	Correct Answer: 3 = Burning coal (Note: While this was accurate for the period of the study, energy sources may have shifted since then.)
	8.	ELS08 - Problem with electric cars:
	•	Correct Answer: 1 = Most electricity is currently produced from fossil fuels (coal, oil, natural gas)

```{r}
# Dewaters and Powers 2011. Assesses participants knowledge of energy names, use, and sources
# - DeWaters, J. E., & Powers, S. E. (2011). Energy literacy of secondary students in New York State (USA): A measure of knowledge, affect, and behavior. Energy Policy, 39(3), 1699–1710. https://doi.org/10.1016/j.enpol.2010.12.049

s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))
s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  rename(id2=ResponseId) |>
  mutate(id=row_number()) |> relocate(id) |> 
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

els <- s1 |> select(id,id2,calc,refClass,state1,state2, ELS01:ELS08)

els2 <- els |> 
  pivot_longer(cols = ELS01:ELS08, names_to = "question", values_to = "response") |> 
  # Convert response to numeric to avoid label issues
  mutate(response = as.numeric(response)) |>
  mutate(question = case_when(
    question == "ELS01" ~ "Electrical energy units (kWh)",
    question == "ELS02" ~ "Energy consumed and appliance power rating",
    question == "ELS03" ~ "Incandescent lightbulb conversion",
    question == "ELS04" ~ "Reason to buy energy star appliances",
    question == "ELS05" ~ "Which appliances uses the most energy",
    question == "ELS06" ~ "Which appliance uses the most electricity",
    question == "ELS07" ~ "Which source provides most electricity in the US",
    question == "ELS08" ~ "Problem with electric cars",
    TRUE ~ NA_character_
  )) |> 
  mutate(correct = case_when(
    question == "Electrical energy units (kWh)" & response == 2 ~ "Correct",
    question == "Energy consumed and appliance power rating" & response == 3 ~ "Correct",
    question == "Incandescent lightbulb conversion" & response == 3 ~ "Correct",
    question == "Reason to buy energy star appliances" & response == 3 ~ "Correct",
    question == "Which appliances uses the most energy" & response == 3 ~ "Correct",
    question == "Which appliance uses the most electricity" & response == 2 ~ "Correct",
    question == "Which source provides most electricity in the US" & response == 3 ~ "Correct",
    question == "Problem with electric cars" & response == 1 ~ "Correct",
    TRUE ~ "Incorrect"
  ))

els_id <- els2 |> group_by(id,id2,state1,state2,refClass) |> 
    summarize(els=mean(correct=="Correct"), .groups="drop")

# plot distribution fo answers to each els question. Separate questions by facets, color by correct/incorrect, make facets go in same order as questions
els2 |> 
  ggplot(aes(x=response, fill=correct)) +
  geom_bar(position="dodge", stat="count") +
  facet_wrap(~question, scales="free_y", ncol=2) +
  scale_fill_manual(values=c("Correct"="green","Incorrect"="red")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title="Energy Literacy Survey", x="Response", y="Count") +
  theme(legend.position = "none")

ggplot(els2, aes(x = as.factor(response), fill = correct)) +
  geom_bar(position = "dodge") +
  facet_wrap(~question, ncol = 2, scales = "free_x") + # Facet by question, 2 columns, free x-axis scales
  scale_fill_manual(values = c("Correct" = "#00ba38", "Incorrect" = "#f8766d")) + # Custom colors
  labs(
    title = "Distribution of Answers to Energy Literacy Survey Questions",
    subtitle = "DeWaters & Powers (2011)",
    x = "Response Option",
    y = "Number of Respondents",
    fill = "Correctness"
  ) +
  theme_bw() + # Use a clean theme
  theme(
    strip.text = element_text(size = 10, face = "bold"), # Adjust facet label appearance
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1), # Rotate x-axis labels for readability
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10),
    legend.position = "bottom", # Move legend to the bottom
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

# score per id
els2 |> 
  group_by(id) |> 
  summarise(score=sum(correct=="Correct")/n(), .groups="drop") |> 
  ggplot(aes(x=score)) +
  geom_histogram(binwidth=.1, fill="blue", color="black") +
  labs(title="Energy Literacy Survey Score", x="Score", y="Count") +
  theme_minimal()

# score per id, color by ReferenceClass
els2 |> 
  group_by(id, refClass) |> 
  summarise(score=sum(correct=="Correct")/n(), .groups="drop") |> 
  ggplot(aes(x=score, fill=refClass)) +
  geom_histogram(binwidth=.1, color="black") +
  labs(title="Energy Literacy Survey Score", x="Score", y="Count") +
  theme_minimal()

# compare refClass averages, stat_halfeye
els2 |> 
  group_by(id,refClass) |> 
  summarise(score=sum(correct=="Correct")/n(), .groups="drop") |> 
  ggplot(aes(x=refClass, y=score, col=refClass)) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
    labs(title="Energy Literacy Survey Score", x="Reference Class", y="Score")

els2 |>
  group_by(refClass, question) |>
  summarize(mean_correct = mean(correct == "Correct", na.rm = TRUE)) |>
  ggplot(aes(x = question, y = mean_correct, fill = refClass)) +
  geom_col(position = "dodge") +
  labs(title = "Mean Correctness by Reference Class and Question",
       x = "Question",
       y = "Proportion Correct",
       fill = "Reference Class") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

r$> els_id
# A tibble: 252 × 6
      id id2               state1   state2 refClass     els
   <int> <chr>             <chr>    <chr>  <chr>      <dbl>
 1     1 R_2QMsYuHS385pTj1 Colorado Texas  Percentage 0.125
 2     2 R_2tzIyZavpl1RZBY Colorado Texas  Percentage 0.75 
 3     3 R_3Gm1Y4QzSlyrEWo Colorado Texas  Percentage 0.75 
 4     4 R_Djh2q728ibyNWjD Colorado Texas  Percentage 0.875
 5     5 R_3fAXNbIZ7kFieSu Colorado Texas  Percentage 0.125
 6     6 R_1JECzl10KgckJXg Colorado Texas  Percentage 0.375
 7     7 R_C2IQAZim25sij4d Colorado Texas  Percentage 0.25 
 8     8 R_5bbi3tpMOhuGBIR Colorado Texas  Percentage 0.75 
 9     9 R_2BsVSAOpehXW9Cr Colorado Texas  Percentage 0.25 
10    10 R_325F8wRln52w1QN Colorado Texas  Percentage 0.625

head(s1_new |> filter(appliance=="Cooling",block==1,plan=="plan1"),n=16)
# A tibble: 16 × 22
# Groups:   id [16]
      id id2           state1 state2 refClass calc  edu   modality state Duration__in_seconds_ appliance   els family state_avg plan  value block pct_goal change state_p_dif state_f_dif less_avg
   <int> <chr>         <chr>  <chr>  <chr>    <chr> <fct> <chr>    <chr>                 <dbl> <chr>     <dbl>  <dbl> <dbl+lbl> <fct> <dbl> <dbl>    <dbl>  <dbl>       <dbl>       <dbl> <lgl>   
 1     1 R_2QMsYuHS38… Color… Texas  Percent… Used… Grad… P        COL                    1008 Cooling   0.125    697  498      plan1   600     1     0.15    -97         102         199 FALSE   
 2     2 R_2tzIyZavpl… Color… Texas  Percent… Used… Coll… P        COL                    1126 Cooling   0.75     697  498      plan1   500     1     0.15   -197           2         199 FALSE   
 3     3 R_3Gm1Y4QzSl… Color… Texas  Percent… Used… Coll… P        COL                     870 Cooling   0.75     697  498      plan1   592     1     0.15   -105          94         199 FALSE   
 4     4 R_Djh2q728ib… Color… Texas  Percent… Used… Coll… P        COL                    2763 Cooling   0.875    697  498      plan1   400     1     0.15   -297         -98         199 TRUE    
 5     5 R_3fAXNbIZ7k… Color… Texas  Percent… No C… Grad… P        COL                     920 Cooling   0.125    697  498      plan1   320     1     0.15   -377        -178         199 TRUE    
 6     6 R_1JECzl10Kg… Color… Texas  Percent… No C… Grad… P        COL                     878 Cooling   0.375    697  498      plan1   550     1     0.15   -147          52         199 FALSE   
 7     7 R_C2IQAZim25… Color… Texas  Percent… Used… Grad… P        COL                    2680 Cooling   0.25     697  498      plan1   500     1     0.15   -197           2         199 FALSE   
 8     8 R_5bbi3tpMOh… Color… Texas  Percent… Used… Some… P        COL                    1628 Cooling   0.75     697  498      plan1   500     1     0.15   -197           2         199 FALSE   
 9     9 R_2BsVSAOpeh… Color… Texas  Percent… Used… Grad… P        COL                     724 Cooling   0.25     697  498      plan1   650     1     0.15    -47         152         199 FALSE   
10    10 R_325F8wRln5… Color… Texas  Percent… Used… Some… P        COL                     913 Cooling   0.625    697  498      plan1   593     1     0.15   -104          95         199 FALSE

```{r}

library(dplyr)
library(tidyr)
library(corrplot)

# Create a wide dataframe where each column is a question, and the values are 1 (Correct) or 0 (Incorrect)
els_wide <- els2 |>
  mutate(correct_binary = ifelse(correct == "Correct", 1, 0)) |>
  # mutate number correct for each sbj, then filter out those with less than 5, 
    group_by(id) |>
    mutate(n_correct = sum(correct_binary)) |>
    filter(n_correct >= 2) |>
    ungroup() |>
  pivot_wider(id_cols = c(id), names_from = question, values_from = correct_binary) |>
  select(-id) # Remove the ID column for correlation analysis

# Calculate the correlation matrix
cor_matrix <- cor(els_wide, use = "pairwise.complete.obs") # Handle missing values

# Visualize the correlation matrix using corrplot
corrplot(cor_matrix, 
         method = "color",  # Use colored squares to represent correlations
         type = "upper",    # Show only the upper triangle
         tl.col = "black",  # Color of text labels
         tl.srt = 45,       # Rotate text labels for better readability
         addCoef.col = "black", # Add correlation coefficients to the plot
         number.cex = 0.7,   # Adjust the size of the correlation coefficients
         diag = FALSE       # Exclude the diagonal (which is always 1)
)

els2 |>
  group_by(refClass, question) |>
  summarize(mean_correct = mean(correct == "Correct", na.rm = TRUE)) |>
  ggplot(aes(x = question, y = mean_correct, fill = refClass)) +
  geom_col(position = "dodge") +
  labs(title = "Mean Correctness by Reference Class and Question",
       x = "Question",
       y = "Proportion Correct",
       fill = "Reference Class") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

els2 |>
  group_by(calc, question) |>
  summarize(mean_correct = mean(correct == "Correct", na.rm = TRUE)) |>
  ggplot(aes(x = question, y = mean_correct, fill = calc)) +
  geom_col(position = "dodge") +
  labs(title = "Mean Correctness by Calculator Use",
       x = "Question",
       y = "Proportion Correct",
       fill = "Calculator Use") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

library(factoextra)

# Perform PCA
pca_result <- prcomp(els_wide, scale. = TRUE)

# Visualize the results
fviz_pca_var(pca_result, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE)

```