### 2025-01-12 assets/figure2.tex
updated methods and appendix.

\documentclass{article}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}

\begin{document}

\begin{figure}[h]
\centering
\caption{Example energy planning task trials across conditions. Participants were asked to allocate energy usage to meet a 15\% reduction goal. The format of the reduction goal was manipulated to be either kilowatt hours (kWh), a percentage (\%), or U.S. dollars (USD).}

\begin{minipage}{0.48\textwidth}
\centering
\subcaption{kWh Condition}
\label{fig:kwh}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Category} & \textbf{\begin{tabular}[c]{@{}c@{}}Electricity Used\\ Last Year by\\ the Wells Family\\ (kWh)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Average Electricity\\ Used Last Year by\\ Households in\\ Colorado (kWh)\end{tabular}} & \textbf{Action Plan} \\ \midrule
Cooling (Central A/C) & 697 & 498 &  \\
Heating the Home & 18,052 & 16,411 &  \\
Water Heating & 11,667 & 5,832 &  \\
Refrigerator & 1,370 & 1,142 &  \\
\begin{tabular}[c]{@{}l@{}}Other (Television,\\ Lighting, Electronics,\\ Washer/Dryer, etc.)\end{tabular} & 7,982 & 6,652 &  \\ \midrule
Total kWh & 39,768 & 30,535 &  \\ \midrule
\textbf{Target Reduction} & \multicolumn{3}{c}{\textbf{5,965 kWh}} \\ \bottomrule
\end{tabular}%
}
\end{minipage}\hfill
\begin{minipage}{0.48\textwidth}
\centering
\subcaption{Percentage Condition}
\label{fig:percent}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Category} & \textbf{\begin{tabular}[c]{@{}c@{}}Electricity Used\\ Last Year by\\ the Wells Family\\ (kWh)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Average Electricity\\ Used Last Year by\\ Households in\\ Colorado (kWh)\end{tabular}} & \textbf{Action Plan} \\ \midrule
Cooling (Central A/C) & 697 & 498 &  \\
Heating the Home & 18,052 & 16,411 &  \\
Water Heating & 11,667 & 5,832 &  \\
Refrigerator & 1,370 & 1,142 &  \\
\begin{tabular}[c]{@{}l@{}}Other (Television,\\ Lighting, Electronics,\\ Washer/Dryer, etc.)\end{tabular} & 7,982 & 6,652 &  \\ \midrule
Total kWh & 39,768 & 30,535 &  \\ \midrule
\textbf{Target Reduction} & \multicolumn{3}{c}{\textbf{15\%}} \\ \bottomrule
\end{tabular}%
}
\end{minipage}

\bigskip % Add some space between the tables

\begin{minipage}{0.48\textwidth}
\centering
\subcaption{USD Condition}
\label{fig:usd}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Category} & \textbf{\begin{tabular}[c]{@{}c@{}}Electricity Used\\ Last Year by\\ the Wells Family\\ (kWh)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Average Electricity\\ Used Last Year by\\ Households in\\ Colorado (kWh)\end{tabular}} & \textbf{Action Plan} \\ \midrule
Cooling (Central A/C) & 697 & 498 &  \\
Heating the Home & 18,052 & 16,411 &  \\
Water Heating & 11,667 & 5,832 &  \\
Refrigerator & 1,370 & 1,142 &  \\
\begin{tabular}[c]{@{}l@{}}Other (Television,\\ Lighting, Electronics,\\ Washer/Dryer, etc.)\end{tabular} & 7,982 & 6,652 &  \\ \midrule
Total kWh & 39,768 & 30,535 &  \\ \midrule
\textbf{Target Reduction} & \multicolumn{3}{c}{\textbf{\$656}} \\ \bottomrule
\end{tabular}%
}
\end{minipage}
\end{figure}

\end{document}