### 2025-04-09 poster/after-body.tex
code to generate poster for mpa

% This part runs after the main document content

\end{multicols} % End the 3-column layout

% Draw a simple footer line with a slightly taller background
\AddToHook{shipout/background}{% Add to background hook for each page
  \begin{tikzpicture}[remember picture, overlay]
    % Extend footer area height slightly
    \fill[purduegold] (current page.south west) rectangle ([yshift=0.53in]current page.south east);

    % Adjust Text position in the footer
    \node[anchor=south east, xshift=-0.75in, yshift=0.2in, % Shift down slightly
          text=darkgray, font=\small\sffamily] % Text color and font
          at (current page.south east) 
          {Communication and Cognition Lab | Purdue University};

    % Adjust email position in the footer
    \node[anchor=south west, xshift=0.75in, yshift=0.05in, % Shift down slightly
          text=darkgray, font=\small\sffamily] % Text color and font
          at (current page.south west) 
          {\textsuperscript{*}Corresponding author: \href{mailto:tegorman@purdue.edu}{tegorman@purdue.edu}};
  \end{tikzpicture}
}

% \end{multicols} % End the 3-column layout

% % Draw background color and text for footer area using TikZ overlay
% \AddToHook{shipout/background}{% Add to background hook for each page
%   \begin{tikzpicture}[remember picture, overlay]
%     % Fill footer area with Purdue Gold - adjust height (1in) as needed
%     \fill[purduegold] (current page.south west) rectangle ([yshift=0.8in]current page.south east);

%     % Add Text to the footer - Adjust position and content
%     \node[anchor=south east, xshift=-0.75in, yshift=0.25in, % Position from bottom right
%           text=darkgray, font=\small\sffamily] % Text color and font
%           at (current page.south east) 
%           {Communication and Cognition Lab | Purdue University};

%     % Add email to the left side of the footer
%     \node[anchor=south west, xshift=0.75in, yshift=0.25in, % Position from bottom left
%           text=darkgray, font=\small\sffamily] % Text color and font
%           at (current page.south west) 
%           {\textsuperscript{*}Corresponding author: \href{mailto:tegorman@purdue.edu}{tegorman@purdue.edu}};

%     % Optional: Add a logo - Uncomment and adjust path/size
%     % \node[anchor=south west, xshift=0.75in, yshift=0.15in] 
%     %       at (current page.south west) {
%     %   \includegraphics[height=0.5in]{path/to/purdue_logo.png} % Adjust height/path
%     % };
%   \end{tikzpicture}
% }