### 2025-04-09 poster/before-body.tex

% before-body.tex

% Replace the current header background with a more subtle approach
\AddToHook{shipout/background}{% Add to background hook for each page
  \begin{tikzpicture}[remember picture, overlay]
    % Fill header area with a much thinner line instead of full background
    \fill[purduegold] (current page.north west) rectangle ([yshift=-0.2in]current page.north east);
  \end{tikzpicture}
}

% --- Logo and Title Block Setup ---
\noindent % Prevent potential paragraph indentation
\makebox[\linewidth]{% Create a box spanning the full line width
  \begin{minipage}[t]{0.15\linewidth} % Left Logo Column
    \raggedright
    \vspace*{0pt} % Align content to the top
    % First logo (Purdue logo)
    \includegraphics[width=0.80\linewidth]{assets/PU-H-Full-RGB.png}

    \vspace{0.12in} % Space between logos

    % Second logo (CLA logo)
    \includegraphics[width=0.75\linewidth]{assets/CLA_logo1.png}
  \end{minipage}%
  \hfill
  \begin{minipage}[t]{0.65\linewidth} % Center Text Column
    \centering
    \vspace*{0pt} % Align content to the top

    % --- Title ---
    {\color{darkgray}\Huge\bfseries\sffamily Planning to Save Energy: How Information Format Affects Accuracy\par}
    \vspace{0.15in} % Space after title

    % --- Authors ---
    {\color{purduegray}\Large\sffamily
    Thomas E. Gorman\textsuperscript{1,2,*},
    Torsten Reimer\textsuperscript{1,2},
    Juan Pablo Loaiza Ramirez\textsuperscript{1,2},
    Hayden Barber\textsuperscript{3}
    \par}
    \vspace{0.08in} % Space after authors

    % --- Affiliations ---
    {\color{purduegray}\large\sffamily
    \textsuperscript{1}Communication and Cognition Lab, Purdue University, West Lafayette, IN, USA \\
    \textsuperscript{2}College of Liberal Arts Research Academy, Purdue University, West Lafayette, IN, USA \\
    \textsuperscript{3}School of Communication \& Journalism, South Dakota State University, Brookings, SD, USA \\
    \vspace{0.05in} % Space after affiliations
    \par}

  \end{minipage}%
  \hfill
  \begin{minipage}[t]{0.10\linewidth} % Right Logo Column
    \raggedleft
    \vspace*{0pt} % Adjust slightly if needed e.g., \vspace*{-2pt}
    \includegraphics[scale=0.42]{assets/nsf_logo1.png}
  \end{minipage}%
} % End \makebox

% Adjust space AFTER the header block, before the multicolumn content starts.
\vspace{0.5in} % Or adjust as needed

% Start the 3-column layout
\begin{multicols}{3}
\raggedright % Justify text to the left, better for narrow columns