### 2025-04-09 poster/poster.qmd

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
---

## Introduction

*   **Problem:** Energy insecurity is a critical public health concern, especially among low-income households who face difficult choices and unsafe coping strategies [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021]. Furthermore, residential energy use significantly contributes to climate change, necessitating sustainable solutions [@farghaliStrategiesEnergyContext2023].
*   **Gap:** While interventions exist, success hinges on communication format [@fischerFeedbackHouseholdElectricity2008]. Specifically, how do different **numerical representations (kWh, %, USD)** of energy reduction goals influence consumers' ability to create *accurate*, specific conservation plans for appliance use?
*   **Concepts & Theory:** This study draws on principles of energy literacy [@dewatersEnergyLiteracySecondary2011], reference class effects (absolute units like kWh may simplify calculations compared to relative units like % or derived units like USD) [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015], and the importance of procedural knowledge (knowing *how* to act on a goal) [@tonkeProvidingProceduralKnowledge2024].

### Research Questions & Hypotheses

*   **Primary Question:** How does the format (kWh, %, USD) used to present an energy reduction goal influence the accuracy of household energy conservation planning?
*   **H1 (Reference Class):** Presenting goals in **absolute units (kWh)** will lead to higher planning accuracy compared to percentage (%) or monetary (USD) formats. 
*   **H2 (Energy Literacy):** Individuals with **higher energy literacy** will demonstrate more accurate planning, regardless of format.

## Methods

*   **Design:** Two online experiments (N=229 & N=190 via MTurk) using a simulated household energy planning task.
*   **Task:** Participants adjusted appliance energy use (kWh) for hypothetical families to meet a specified reduction goal.
*   **Manipulation:** Goal presentation format (Between-Subjects): **kWh** vs. **%** vs. **USD**.
*   **Exp 2 Factors:** Also varied Goal Difficulty (10/15%) & Numerical Rounding (exact/rounded).

### Figure 1: Procedure and Task Example

<!-- ![Participants were shown a family's previous year usage by appliance alongside state averages. They were asked to create two action plans by adjusting the 'kWh next year' for each appliance to meet a specific total reduction goal. The goal was presented differently based on condition: as a percentage (15%), absolute kWh (reduce by 5965 kWh), or monetary savings ($656, with conversion rate provided).](assets/images/wells10e_pub.png){width="70%" fig-align="center"} -->

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.28\linewidth}
    \centering
    \subcaptionbox{Procedure overview \label{fig-mermaid}}[0.95\linewidth]{%
      \includegraphics[width=\linewidth, height=9.5in,keepaspectratio]{assets/images/mermaid_proc.png}
    }
  \end{minipage}
  \hspace{0.5em}
  \begin{minipage}[t]{0.68\linewidth}
    \centering
    \subcaptionbox{Example of planning task interface \label{fig-task}}[0.95\linewidth]{%
      \includegraphics[width=\linewidth,height=9.5in, keepaspectratio]{assets/images/tab.png}
    }
  \end{minipage}
}

\end{figure}
```

\columnbreak

## Results

*  **Key Measures:**
    *   **Planning Accuracy:** Absolute deviation between plan reduction and the goal reduction (ordinal bins: exact, minor, large error).
    *   **Energy Literacy:** 8-item knowledge scale score [@dewatersEnergyLiteracySecondary2011].
*  **Analysis:** Bayesian mixed-effects regression (Ordinal & Linear) controlling for participant and scenario random effects: $\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})$
* **Goal Format (H1 Supported):** Planning accuracy was significantly higher when the energy reduction goal was presented in absolute units (kWh) compared to percentages (%) or U.S. dollars (USD) across both experiments. 
    * *Exp 1:* USD (*OR* = 22.9) and % (*OR* = 4.2) formats had significantly higher odds of lower accuracy than kWh.
    * *Exp 2:* Similar significant effect for USD (*OR* = 9.68); the trend persisted for % (*OR* = 2.78), though the CI included 1.0. USD/percentage formats reduced the probability of exact matches and increased large errors compared to kWh.
* **Energy Literacy (H2 Supported):** Higher energy literacy scores were significantly associated with more accurate planning (lower error) in both experiments (Exp 1: $B$ = -2.35; Exp 2: $B$  = -3.21). (See Fig 3b).
* **Exploratory Findings (Exp 2):**
    * Using rounded numbers for prior usage slightly improved accuracy (Est. = -0.53). No effect of Goal Difficulty.

### Figure 2: Reference Class Effect on Planning Error (Exp 1)

![Distribution of log absolute error by goal format condition (Exp 1). Lower values indicate higher accuracy.](assets/images/fig-s1-log-dist-1.png){width="80%"}

### Figure 3

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.44\linewidth}
    \centering
    \subcaptionbox{Posterior predictive check for the ordinal regression model. Blue bars show the observed frequencies per accuracy level, dots represent the model predictions.\label{fig-s1-ppd}}[0.93\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s1-ppd-1.png}
    }
  \end{minipage}
  \hspace{0.01em}
  \begin{minipage}[t]{0.44\linewidth}
  \centering
    \subcaptionbox{Relationship between energy literacy score and log absolute error, controlling for random effects \label{fig-s1-els}}[0.93\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s1-els-1.png}
    }
  \end{minipage}
}

\end{figure}
```
\columnbreak

### Figure 4: Reference Class Effect on Planning Error (Exp 2)

![Distribution of log absolute error by goal format condition (Exp 2). Lower values indicate higher accuracy.](assets/images/fig-s2-log-dist-1.png){width="83%"}

<!-- ### Figure 5

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.44\linewidth}
    \centering
    \subcaptionbox{Posterior predictive check for the Bayesian ordinal regression model. Blue bars show the observed frequencies for each accuracy level, dots with error bars represent the model predictions.\label{fig-s2-ppd}}[0.90\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s2-ppd-1.png}
    }
  \end{minipage}
  \hspace{0.01em}
  \begin{minipage}[t]{0.44\linewidth}
  \centering
    \subcaptionbox{Relationship between energy literacy score and log absolute error, controlling for random effects of participant and family scenario \label{fig-s2-els}}[0.90\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s2-els-1.png}
    }
  \end{minipage}
}

\end{figure}
``` -->

## Discussion

*   **Key Finding:** Goal format significantly impacts planning accuracy. **Presenting goals in absolute kWh is superior** to relative (%) or monetary (USD) formats, likely by simplifying calculations and reducing cognitive load [@gigerenzerSimpleToolsUnderstanding2003].
*   **Energy Literacy Matters:** Higher domain knowledge consistently predicted better planning accuracy, emphasizing the role of energy education [@canfieldPerceptionsElectricityuseCommunications2017].
*   **Limitations:** Lab simulation, short duration, self-reported calculator use.
*   **Future Directions:** Explore combining kWh format with actionable tips [@tonkeProvidingProceduralKnowledge2024]; test interactive tools/visualizations.

## References

::: {#refs}
:::