### 2025-04-09 poster/poster.qmd
update poster

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
---

## Introduction

*   **Problem:** Energy insecurity is a critical public health concern [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021]. 
*   **Prior Work:** While interventions exist, success often hinges on communication format [@canfieldPerceptionsElectricityuseCommunications2017; @fischerFeedbackHouseholdElectricity2008], such as the use of tables vs. figures, the framing of information (e.g., savings vs. costs), and energy literacy. Many reasoning tasks involving quantities are also influenced by reference class (e.g., absolute units like kWh may simplify calculations compared to relative units like % or derived units like USD) [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].

### Research Questions & Hypotheses

*   **Primary Question:** How does the Reference class (kWh, %, USD) used to present a reduction goal influence the accuracy of energy planning?
*   **H1 (Reference Class):** Goals presented in **absolute units (kWh)** will benefit planning accuracy compared to percentage (%) or monetary (USD) formats. 
*   **H2 (Energy Literacy):** Individuals with **higher energy literacy** will demonstrate more accurate planning, regardless of format.

## Methods

*   **Design:** Two online experiments (N=229 & N=190 via MTurk) using a simulated household energy planning task.
*   **Planning Task:** Propose appliance energy use (kWh) for hypothetical families to meet a target reduction goal.
*   **Manipulation:** Reference Class of reduction goal (Between-Subjects): **kWh** vs. **%** vs. **USD**.
*   **Exp 2 Factors:** Goal Difficulty (10% vs. 15% reduction) and Numerical Rounding (exact vs. rounded).

### Figure 1: Procedure and Task Example

<!-- ![Participants were shown a family's previous year usage by appliance alongside state averages. They were asked to create two action plans by adjusting the 'kWh next year' for each appliance to meet a specific total reduction goal. The goal was presented differently based on condition: as a percentage (15%), absolute kWh (reduce by 5965 kWh), or monetary savings ($656, with conversion rate provided).](assets/images/wells10e_pub.png){width="70%" fig-align="center"} -->

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.28\linewidth}
    \centering
    \subcaptionbox{Procedure overview \label{fig-mermaid}}[0.95\linewidth]{%
      \includegraphics[width=\linewidth, height=9.5in,keepaspectratio]{assets/images/mermaid_proc.png}
    }
  \end{minipage}
  \hspace{0.5em}
  \begin{minipage}[t]{0.68\linewidth}
    \centering
    \subcaptionbox{Example of planning task interface \label{fig-task}}[0.95\linewidth]{%
      \includegraphics[width=\linewidth,height=9.5in, keepaspectratio]{assets/images/tab.png}
    }
  \end{minipage}
}

\end{figure}
```

\columnbreak

## Results

*  **Key Measures:**
    *   **Planning Accuracy:**  Deviation between intended reduction goal and proposed plan (ordinal bins: exact, minor, large error).
    *   **Energy Literacy:** 8-item knowledge scale score [@dewatersEnergyLiteracySecondary2011].
*  **Analysis:** Bayesian mixed-effects regression: $\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})$
* **Effect of reference class (H1 Supported):** Smallest planning errors when reduction goal was presented in absolute units (kWh) compared to percentages (%) or dollars (USD) across both experiments. 
* **Energy Literacy (H2 Supported):** Higher energy literacy scores associated with more accurate planning (lower error) in both experiments.

### Figure 2: Reference Class Effect on Planning Error (Exp 1)

![Distribution of log absolute error by goal format condition (Exp 1). Lower values indicate higher accuracy.](assets/images/fig-s1-log-dist-1.png){width="85%"}

### Figure 3 - Bayesian Regression Results (Exp 1)

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.44\linewidth}
    \centering
    \subcaptionbox{PPC for ordinal model. Blue bars are observed frequencies per accuracy level, dots show model predictions.\label{fig-s1-ppd}}[0.93\linewidth]{%
    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s1-ppd-1.png}
    }
  \end{minipage}
  \hspace{0.01em}
  \begin{minipage}[t]{0.44\linewidth}
  \centering
    \subcaptionbox{Relationship between energy literacy score and log absolute error of reduction plans \label{fig-s1-els}}[0.93\linewidth]{%
    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s1-els-1.png}
    }
  \end{minipage}
}

\end{figure}
```
\columnbreak

### Figure 4: Reference Class Effect on Planning Error (Exp 2)

![Distribution of log absolute error by goal format condition (Exp 2). Lower values indicate higher accuracy.](assets/images/fig-s2-log-dist-1.png){width="85%"}

### Figure 5 - Bayesian Regression Results (Exp 2)

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.44\linewidth}
    \centering
    \subcaptionbox{PPC for ordinal model. Blue bars are observed frequencies per accuracy level, dots show model predictions.\label{fig-s2-ppd}}[0.90\linewidth]{%
    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s2-ppd-1.png}
    }
  \end{minipage}
  \hspace{0.01em}
  \begin{minipage}[t]{0.44\linewidth}
  \centering
    \subcaptionbox{Relationship between energy literacy score and log absolute error of reduction plans. \label{fig-s2-els}}[0.90\linewidth]{%
    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s2-els-1.png}
    }
  \end{minipage}
}

\end{figure}
```

## Discussion

*   **Key Finding:** Reference class  significantly impacts planning accuracy. **Presenting goals in absolute kWh is superior** to relative (%) or monetary (USD) formats, likely by simplifying calculations and reducing cognitive load.
*   **Energy Literacy Matters:** Higher domain knowledge consistently predicted better planning accuracy, emphasizing the role of energy education.
*   **Limitations:** Lab simulation, short duration, self-reported calculator use.

## References

::: {#refs}
:::